### 2022-03-14 manuscript.Rmd
First commit.

---
title: "On the use of relative brain size"
author: "Simeon Q. Smeele $^{1,2,3}$"
date: "March 2022"
output:
  pdf_document: default
header-includes:
- \usepackage{lineno}
- \linenumbers
editor_options:
  chunk_output_type: console
bibliography: bibliography.bib
---

```{r include=FALSE}
# Libraries
## cmdstanr can be installed from Cran, but you need to run 
## library(cmstanr) and install_cmdstan() to finish installing 
library(cmdstanr)
library(scales)

# Paths
path_response_stan = 'response.stan'
path_predictor_stan = 'predictor.stan'
path_sem_stan = 'sem.stan'
path_pdf_response = 'response.pdf'
path_pdf_predictor = 'predictor.pdf'
```

$^1$Cognitive & Cultural Ecology Research Group, Max Planck Institute of Animal Behavior, Radolfzell, Germany

$^2$Department of Human Behavior, Ecology and Culture, Max Planck Institute for Evolutionary
Anthropology, Leipzig, Germany

$^3$Department of Biology, University of Konstanz, Konstanz, Germany

**Subject areas:** evolution

**Keywords:**: relative brain size, comparative analysis, structural equation model, Bayesian

**Author for correspondence:** Simeon Q. Smeele, ssmeele@ab.mpg.de

# Abstract

There is a long standing interest in the effect of relative brain size on other life history variables in a comparative context. Historically, residuals have been used to calculate these effects, but more recently it has been recognised that regression on residuals is not good practice. Instead, absolute brain size and body size are included in a single regression, with the idea that this controls for allometry. This practice leads to a second, potentially even greater estimation error when both relative brain size and body size are a cause of the third variable in question. In this study I use a simple simulation to show how the effect of body mass is masked if absolute brain size is included. I propose the use of structural equation models to simultaneously estimate relative brain size and its effect on the third variable, leaving the effect of body mass unconfounded.

# Introduction

Brain size is often theorised to be important in the life history of species and has been linked to everything from innovation rates [@lefebvre2004brains], to sociality [@dunbar2007evolution] and longevity [@minias2017longevity]. It is important to note that it is often the *relative* and not the *absolute* size of the brain that is of interest. When talking about relative brain size, it is always implicit that brain size is relative to body size or a specific brain region relative to the whole brain. This to control for allometric scaling. There exists a large diversity of relative brain size measures [@healy2007critique], but some decades ago the use of residual brain size was proposed and quickly became the most popular [@jerison1973evolution]. This measure is simply the residual from a model that has log body size as predictor and log brain size as response variable. It is an intuitively attractive approach, since it seems to get at *that extra bit of brain* a species has. Multiple challenges were recognised early on. Probably the most debated is the effect of phylogeny [@armstrong1983relative]. When comparing across multiple species nested at different levels of taxonomy, how does one estimate the 'true' intercept and slope of the model? Several techniques have been proposed to mitigate this, but the debate is ongoing.

More recently another statistical caveat has been highlighted (for an overview see @freckleton2002misuse). The use of residuals is great for visualisation, but cannot be used in statistical models. The proposed solution is to simply include body size as a predictor variable in the final model, which is very similar to the use of residual brain size, if - and only if - relative brain size is the response variable. This approach takes care of both the phylogenetic signal (since models can include the phylogenetic variance-covariance matrix) and correct estimation of uncertainty. However, one major issue has been overlooked: if relative brain size is a predictor variable, including body size as a second predictor can lead to incorrect inference, because body size is not controlling for variation in brain size, as brain size is not the response variable. Brain size does, however, contain information about body size as body size is a cause of brain size. This will result in a reduction of the estimated effect of the body size. To properly include both relative brain size and body size one needs a system that contains regressions with brain size as response (of body size) and as predictor (of the third variable).

A structural equation model is such a system. It contains regressions for each variable and allows brain size to be response and outcome variable simultaneously [@bowen2011structural]. Structural equation models have been used before to include relative brain size in a comparative study in combination with phylogeny, measurement error and imputation [REF own paper]. The aim of this current paper is to show the estimation error in ordinary linear regressions using a simulation and propose a simple Bayesian structural equation model as a solution that can be easily adapted to most comparative studies.

# Methods

To show the difference between a case where relative brain size is the *response* variable and where it is a *predictor* variable I simulated two simple cases with three variables. To avoid discussion about directionality I have simply named the third variable *z*. To further simplify the simulation I have not attempted to simulate realistic values for body size and relative brain size, but just used values with mean = 0 and standard deviation = 1. This allows me to draw general conclusions that are not sensitive to the scale of the variables.

I simulated 20 datasets per case and present parameter estimates for all datasets. Frequentist linear models were fitted with the *lm* function from base R [@R]. Bayesian models were fitted using the *cmstanr* package [@cmdstanr], which runs the No U-turn Sampler in Stan [@gelman2015stan] with four chains and default sampler settings.

## Case I: relative brain size as response variable

In the first case absolute brain size is caused by body size and *z*. The interest of the study is to what extent *z* causes additional increase in brain size. In other words, to what extend *z* correlates with relative brain size. I simulated 20 data sets with 100 species with the following structure:

$$\begin{aligned}
\text{body size} & \sim \text{normal}(0,\ 1) \\
z & \sim \text{normal}(0,\ 1) \\
\text{brain size} & \sim \text{normal}(\mu_{brain},\ 1) \\
\mu_{brain} & = 1 * \text{body size} + 1 * z
\end{aligned}$$

I analysed the resulting data with a frequentist linear model and with the Bayesian equivalent. Then I plotted the estimated effect of body size and *z* to show how well parameters were retrieved.

## Case II: relative brain size as predictor variable

In the second case both body size and relative brain size are a predictor of *z*. The interest of the study is to what extend relative brain size causes *z*. I simulated 20 data sets with 100 species with the following structure:

$$\begin{aligned}
\text{body size} & \sim \text{normal}(0,\ 1) \\
\text{relative brain size} & \sim \text{normal}(0,\ 1) \\
\text{brain size} & = \text{relative brain size} + 1 * \text{body size} \\
z & \sim \text{normal}(\mu_z,\ 1) \\
\mu_z & = 1 * \text{body size} + 1 * \text{relative brain size}
\end{aligned}$$

I analysed the resulting data with both a frequentist and Bayesian linear model where brain size and body size were included as predictor variables. Additionally I analysed the data with a Bayesian structural equation model that included sub-models for all causal paths:

$$\begin{aligned}
\text{body size} & \sim \text{normal}(\alpha_\text{body},\ \sigma_\text{body}) \\
\\
\text{brain size} & \sim \text{normal}(\mu_\text{brain},\ \sigma_\text{brain}) \\
\mu_\text{brain[i]} & = \alpha_\text{brain} + \beta_\text{body} * \text{body size}_\text{[i]} \\
\\
z & \sim \text{normal}(\mu_z,\ \sigma_z) \\
\mu_{z\text{[i]}} & = \alpha_z + \zeta_\text{body} * \text{body size}_\text{[i]} + \zeta_\text{brain} * (\text{brain size}_\text{[i]} - \mu_\text{brain[i]})\\
\\
\alpha_\text{body},\ \alpha_\text{brain},\ \alpha_z  & \sim \text{normal}(0,\ 1)\\
\beta_\text{body},\ \zeta_\text{body},\ \zeta_\text{brain} & \sim \text{normal}(0,\ 1)\\
\sigma_\text{body},\ \sigma_\text{brain},\ \sigma_z & \sim \text{exponential}(1)\\
\end{aligned}$$

The model includes a regression for each variable. Body size is not a function of any variable. Brain size is a function of body size. *z* is a function of body size and relative brain size (where relative brain size is the difference between the actual and expected brain size). The last three lines are the priors for all parameters. Note that the priors for the slopes ($\beta$ and $\zeta$) are set to normal(0, 1), which regularises them slightly and assumes no effect of the predictors. For empirical studies theory might provide more informative priors, which would further increase the accuracy of the model.

# Results

For the Case I, were relative brain size was the response variable, both models estimated the parameters very well (see Figure 1). For Case II, where relative brain size was a predictor variable, the effect of brain size was estimated well by all models, but the effect of body size was only estimated correctly by the structural equation model (see Figure 2). Both the frequentist and Bayesian linear model estimated the effect of body mass to be essentially 0.

```{r response, include=FALSE}
# Settings
set.seed(1)
N = 100
z_effect = 1
body_effect = 1
model_response = cmdstan_model(path_response_stan)

# Simulate data 20 times
out = lapply(1:20, function(i){
  body_size = rnorm(N, 1, 1)
  z = rnorm(N, 0, 1)
  mu_brain = body_effect * body_size + z_effect * z
  brain_size = rnorm(N, mu_brain, 1)

  # Run linear model
  linear_fit = lm(brain_size ~ body_size + z)

  # Run Bayesian linear model
  d = list(N = N, brain_size = brain_size, body_size = body_size, z = z) # save data as list
  bay_linear_fit = model_response$sample(data = d, # run model
                                         seed = 1, 
                                         chains = 4, 
                                         parallel_chains = 4, # runs on four threads
                                         refresh = 2000)
  post_linear_fit = bay_linear_fit$draws(format = 'draws_df') # draw posterior samples

  return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit))
})

# Plot the fit
pdf(path_pdf_response, 6, 5)
par(mfrow = c(2,2), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), xlab = '', yaxt = 'n', main = 'linear model')
axis(4)
mtext('density', 4, 2.5)
mtext('effect z', 1, 2.5)
abline(v = z_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
                                   summary(out[[i]]$linear_fit)$coefficients[3,2])), 
                     pch = 16, cex = 3, col = alpha('#D68910', 0.5))
mtext('z', 2, 3, font = 2, las = 2, adj = 0)
plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
     xlab = '', yaxt = 'n', main = 'Bayesian linear model')
abline(v = z_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_z), pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
axis(4)
mtext('density', 4, 2.5)
mtext('effect z', 1, 2.5)

plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
     xlab = '', yaxt = 'n', main = '')
abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2], 
                                   summary(out[[i]]$linear_fit)$coefficients[2,2])), 
                     pch = 16, cex = 3, col = alpha('#D68910', 0.5))
mtext('body', 2, 3, font = 2, las = 2, adj = 0)
axis(4)
mtext('density', 4, 2.5)
mtext('effect body size', 1, 2.5)
plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
     xlab = '', yaxt = 'n', main = '')
abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
axis(4)
mtext('density', 4, 2.5)
mtext('effect body size', 1, 2.5)
dev.off()
```

```{r plot response, echo=FALSE, out.width = "50%", fig.cap= "Parameter estimates from the linear model and Bayesian linear model with brain size as response variable. Dashed grey line is the true value. Orange density plots are normal distributions based on the mean and SE from the linear model. Purple and green density plots are the posterior distributions for the Bayesian models."}
knitr::include_graphics(path_pdf_response)
```

```{r predictor, include=FALSE}
# Settings
set.seed(1)
N = 100
brain_effect = 1
body_effect = 1
model_predictor = cmdstan_model(path_predictor_stan)
model_sem = cmdstan_model(path_sem_stan)

# Simulate data
out = lapply(1:20, function(i){
  body_size = rnorm(N, 0, 1)
  rel_brain_size = rnorm(N, 0, 1)
  brain_size = 1 * body_size + rel_brain_size
  mu_z = body_effect * body_size + brain_effect * rel_brain_size
  z = rnorm(N, mu_z, 1)

  # Run linear model
  linear_fit = lm(z ~ body_size + brain_size)

  # Run Bayesian linear model
  d = list(N = N, brain_size = brain_size, body_size = body_size, z = z)
  bay_linear_fit = model_predictor$sample(data = d, 
                                          seed = 1, 
                                          chains = 4, 
                                          parallel_chains = 4,
                                          refresh = 2000)
  post_linear_fit = bay_linear_fit$draws(format = 'draws_df')

  # Run Bayesian SEM
  bay_sem_fit = model_sem$sample(data = d, 
                                 seed = 1, 
                                 chains = 4, 
                                 parallel_chains = 4,
                                 refresh = 2000)
  post_sem_fit = bay_sem_fit$draws(format = 'draws_df')

  return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit, post_sem_fit = post_sem_fit))
})

# Plot the fit
pdf(path_pdf_predictor, 9, 5)
par(mfrow = c(2,3), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
     xlab = '', yaxt = 'n', main = 'linear model')
abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
                                   summary(out[[i]]$linear_fit)$coefficients[3,2])), 
                     pch = 16, cex = 3, col = alpha('#D68910', 0.5))
axis(4)
mtext('density', 4, 2.5)
mtext('effect brain size', 1, 2.5)
mtext('brain', 2, 4, font = 2, las = 2, adj = 0)
plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
     xlab = '', yaxt = 'n', main = 'Bayesian linear model')
abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_brain), pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
axis(4)
mtext('density', 4, 2.5)
mtext('effect brain size', 1, 2.5)
plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
     xlab = '', yaxt = 'n', main = 'Bayesian SEM model')
abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_brain_z), pch = 16, cex = 3, col = alpha('#16A085', 0.5))
axis(4)
mtext('density', 4, 2.5)
mtext('effect brain size', 1, 2.5)

plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
     xlab = '', yaxt = 'n', main = '')
abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2],
                                   summary(out[[i]]$linear_fit)$coefficients[2,2])), 
                     pch = 16, cex = 3, col = alpha('#D68910', 0.5))
axis(4)
mtext('density', 4, 2.5)
mtext('effect body size', 1, 2.5)
mtext('body', 2, 4, font = 2, las = 2, adj = 0)
plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
     xlab = '', yaxt = 'n', main = '')
abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
axis(4)
mtext('density', 4, 2.5)
mtext('effect body size', 1, 2.5)
plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
     xlab = '', yaxt = 'n', main = '')
abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_body_z), pch = 16, cex = 3, col = alpha('#16A085', 0.5))
axis(4)
mtext('density', 4, 2.5)
mtext('effect body size', 1, 2.5)
dev.off()
```

```{r plot predictor, echo=FALSE, out.width = "72%", fig.cap= "Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable. Dashed grey line is the true value. Orange density plots are normal distributions based on the mean and SE from the linear model. Purple and green density plots are the posterior distributions for the Bayesian models."}
knitr::include_graphics(path_pdf_predictor)
```

# Discussion

Several recent studies have claimed to study the effect of relative brain size by including both absolute brain size and body size as predictors [@street2017coevolution; @gonzalez2010large; @isler2009why]. These studies also reported the direct effect of body size and sometimes drew conclusions based on the sign of this effect. Since all these studies used some version of a linear model (be it phylogenetic and/or Bayesian), they actually tested the effect of absolute brain size, as including body size only accounts for allometry in the response variable.  The simulations in this paper showed that including body mass as additional predictor to control for allometric scaling of brain size works well if relative brain size is the response variable, but not when relative brain size is a predictor variable. Perhaps counter-intuitively, the effect of brain size was still estimated correctly by all models. It was the body size effect that was biased in the linear models. The use of such models would lead to the incorrect inference that there was no effect of body size at all.

One way to create some intuition about what is going on is that absolute brain size (which was the actual predictor variable included in the linear models) contains information about both body size and relative brain size. In a sense this variable controls for the body size effect already, leaving the direct effect of body size 0. In a case where body size itself does not have an effect on *z*, it does not actually need to be included at all. Using absolute brain size would be fine. The variation in brain size due to allometric scaling would just create noise. Using absolute brain size would of course lead to a less precise estimate of the effect of brain size, so using relative brain size from a structural equation model would still be preferable. By presenting a simple version of such a model, I hope to encourage future work to adopt this practice to ensure that all parameter estimates, and the inference drawn based on these, are statistically sound.

**Data accessibility.** All data is generated in the simulation. Code is publicly available at: https://github.com/simeonqs/On_the_use_of_relative_brain_size.

**Competing interests.** I declare I have no competing interests.

**Funding.** I received funding from the International Max Planck Research School for Quantitative Behaviour, Ecology and Evolution.

**Acknowledgements.** XXX

# References

### 2022-03-15 manuscript.Rmd
Adding references and minor rewrites.

@@
 # Libraries
-## cmdstanr can be installed from Cran, but you need to run 
-## library(cmstanr) and install_cmdstan() to finish installing 
+## for installation instructions see the README
 library(cmdstanr)
@@
 # Paths
-path_response_stan = 'response.stan'
+path_response_stan = 'response.stan' # models
 path_predictor_stan = 'predictor.stan'
 path_sem_stan = 'sem.stan'
-path_pdf_response = 'response.pdf'
+path_pdf_response = 'response.pdf' # pdfs
 path_pdf_predictor = 'predictor.pdf'
@@
 $^3$Department of Biology, University of Konstanz, Konstanz, Germany
+
+**ORCID:** 0000-0003-1001-6615
 
@@
 Multiple challenges were recognised early on.
-Probably the most debated is the effect of phylogeny [@armstrong1983relative].
-When comparing across multiple species nested at different levels of taxonomy, how does one estimate the 'true' intercept and slope of the model?
+Probably the most debated is the effect of phylogeny [@armstrong1983relative; @font2019rethinking; @burger2019allometry].
+When comparing across multiple species nested at different levels of taxonomy, how does one estimate the 'true' slope of the model?
 Several techniques have been proposed to mitigate this, but the debate is ongoing.
@@
 The use of residuals is great for visualisation, but cannot be used in statistical models.
-The proposed solution is to simply include body size as a predictor variable in the final model, which is very similar to the use of residual brain size, if - and only if - relative brain size is the response variable.
-This approach takes care of both the phylogenetic signal (since models can include the phylogenetic variance-covariance matrix) and correct estimation of uncertainty.
-However, one major issue has been overlooked: if relative brain size is a predictor variable, including body size as a second predictor can lead to incorrect inference, because body size is not controlling for variation in brain size, as brain size is not the response variable.
-Brain size does, however, contain information about body size as body size is a cause of brain size.
+The proposed solution is to include body size as a predictor variable in the final model, which is very similar to the use of residual brain size, if - and only if - relative brain size is the response variable.
+This approach takes care of both the phylogenetic signal (since the model can include the phylogenetic variance-covariance matrix) and correct estimation of uncertainty.
+However, one major issue has been overlooked: if relative brain size is a predictor variable, including body size as a second predictor can lead to incorrect inference, because body size is not controlling for variation in brain size, since brain size is not the response variable.
+Brain size does, however, contain information about body size since body size is a cause of brain size.
 This will result in a reduction of the estimated effect of the body size.
@@
 It contains regressions for each variable and allows brain size to be response and outcome variable simultaneously [@bowen2011structural].
-Structural equation models have been used before to include relative brain size in a comparative study in combination with phylogeny, measurement error and imputation [REF own paper].
+Structural equation models have been used before to include relative brain size in a comparative study in combination with phylogeny, measurement error and imputation [@smeele2021coevolution].
 The aim of this current paper is to show the estimation error in ordinary linear regressions using a simulation and propose a simple Bayesian structural equation model as a solution that can be easily adapted to most comparative studies.
@@
 *z* is a function of body size and relative brain size (where relative brain size is the difference between the actual and expected brain size).
-The last three lines are the priors for all parameters.
+Relative brain size in this model is very similar to residual brain size, but since it is computed at each iteration information flows in both directions and measurement error is correclty estimated.
+The last three lines of the model are the priors for all parameters.
 Note that the priors for the slopes ($\beta$ and $\zeta$) are set to normal(0, 1), which regularises them slightly and assumes no effect of the predictors.
@@
 
-For the Case I, were relative brain size was the response variable, both models estimated the parameters very well (see Figure 1).
+For Case I, were relative brain size was the response variable, both models estimated the parameters very well (see Figure 1).
 For Case II, where relative brain size was a predictor variable, the effect of brain size was estimated well by all models, but the effect of body size was only estimated correctly by the structural equation model (see Figure 2).
@@
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple and green density plots are the posterior distributions for the Bayesian models."}
+Purple and green density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_response)
@@
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple and green density plots are the posterior distributions for the Bayesian models."}
+Purple and green density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_predictor)
@@
 Since all these studies used some version of a linear model (be it phylogenetic and/or Bayesian), they actually tested the effect of absolute brain size, as including body size only accounts for allometry in the response variable.
-The simulations in this paper showed that including body mass as additional predictor to control for allometric scaling of brain size works well if relative brain size is the response variable, but not when relative brain size is a predictor variable.
+The simulations in this paper showed that including body mass as additional predictor to control for allometric scaling of brain size works well if relative brain size is the response variable, but not if relative brain size is a predictor variable.
 Perhaps counter-intuitively, the effect of brain size was still estimated correctly by all models.
@@
 Using absolute brain size would of course lead to a less precise estimate of the effect of brain size, so using relative brain size from a structural equation model would still be preferable.
-By presenting a simple version of such a model, I hope to encourage future work to adopt this practice to ensure that all parameter estimates, and the inference drawn based on these, are statistically sound.
+The use of these models are not limited to relative brain size, but can be used for any comparative study in which multiple causal paths are of interest.
 

### 2022-03-16 manuscript.Rmd
Few typos corrected.

@@
 output:
+  word_document: default
   pdf_document: default
@@
 
-**Keywords:**: relative brain size, comparative analysis, structural equation model, Bayesian
+**Keywords:** relative brain size, comparative analysis, structural equation model, Bayesian
 
@@
 
-For Case I, were relative brain size was the response variable, both models estimated the parameters very well (see Figure 1).
+For Case I, where relative brain size was the response variable, both models estimated the parameters very well (see Figure 1).
 For Case II, where relative brain size was a predictor variable, the effect of brain size was estimated well by all models, but the effect of body size was only estimated correctly by the structural equation model (see Figure 2).
@@
 Using absolute brain size would of course lead to a less precise estimate of the effect of brain size, so using relative brain size from a structural equation model would still be preferable.
-The use of these models are not limited to relative brain size, but can be used for any comparative study in which multiple causal paths are of interest.
+The use of these models is not limited to relative brain size, but can be used for any comparative study in which multiple causal paths are of interest.
 

### 2022-04-08 supplemental materials.Rmd
First versions. Adding supplemental materials.

---
title: "On the use of relative brain size - supplemental materials"
author: "Simeon Q. Smeele $^{1,2,3}$"
date: "April 2022"
output:
  pdf_document: default
header-includes:
- \usepackage{lineno}
- \linenumbers
editor_options:
  chunk_output_type: console
bibliography: bibliography.bib
---

```{r set-up, include=FALSE}
# Libraries
## for installation instructions see the README
library(cmdstanr)
library(scales)

# Paths
path_response_stan = 'models/response.stan' # models
path_predictor_stan = 'models/predictor.stan'
path_sem_stan = 'models/sem.stan'
path_supplemental_models = 'supplemental models' # folder to supplemental models
path_supplemental_figures = 'supplemental figures' # folder to plot supplemental figures
path_pdf_response_20 = 'supplemental figures/response 20.pdf' # pdfs
path_pdf_response_1000 = 'supplemental figures/response 1000.pdf'
path_pdf_predictor_20 = 'supplemental figures/predictor 20.pdf'
path_pdf_predictor_1000 = 'supplemental figures/predictor 1000.pdf'
path_pdf_response_informative = 'supplemental figures/response informative.pdf'
path_pdf_response_vague = 'supplemental figures/response vague.pdf'
path_pdf_predictor_informative = 'supplemental figures/predictor informative.pdf'
path_pdf_predictor_vague = 'supplemental figures/predictor vague.pdf'
```

$^1$Cognitive & Cultural Ecology Research Group, Max Planck Institute of Animal Behavior, Radolfzell, Germany

$^2$Department of Human Behavior, Ecology and Culture, Max Planck Institute for Evolutionary
Anthropology, Leipzig, Germany

$^3$Department of Biology, University of Konstanz, Konstanz, Germany

# Methods

## Sample size

The main text contained results for data sets with 100 species. To see how the models perform with a smaller or larger dataset, I simulated data for 20 and 1000 species, with 20 simulations per case and sample size.

## Priors

The main text contained results for Bayesian models with slightly regularising priors. In emperical data sets, one can often choose more informative priors. To test the effect of this I analysed the dataset with 100 species with priors that only allow for positive slopes, with the mean set to the simulated value (1). I also restricted the priors for the intecept parameters to normal(0, 0.25).

To further test the robustness of the models I analysed the data with vague priors, settings intercept and slope to normal(0, 10) and standard variation parameters to exponential(0.1) (note that the mean for the exponential distribution is the inverse of the rate, such that the mean standard deviation here is 10).

# Results

## Sample size

### Case I: relative brain size as response variable

```{r response-sampsize, include=FALSE}
for(N in c(20, 1000)){
  # Settings
  set.seed(1)
  z_effect = 1
  body_effect = 1
  model_response = cmdstan_model(path_response_stan)
  if(N == 20) ylim = c(0, 5) else ylim = c(0, 15)

  # Simulate data 20 times
  out = lapply(1:20, function(i){
    body_size = rnorm(N, 1, 1)
    z = rnorm(N, 0, 1)
    mu_brain = body_effect * body_size + z_effect * z
    brain_size = rnorm(N, mu_brain, 1)

    # Run linear model
    linear_fit = lm(brain_size ~ body_size + z)

    # Run Bayesian linear model
    d = list(N = N, brain_size = brain_size, body_size = body_size, z = z) # save data as list
    bay_linear_fit = model_response$sample(data = d, # run model
                                           seed = 1, 
                                           chains = 4, 
                                           parallel_chains = 4, # runs on four threads
                                           refresh = 2000)
    post_linear_fit = bay_linear_fit$draws(format = 'draws_df') # draw posterior samples

    return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit))
  })

  # Plot the fit
  pdf(sprintf('%s/response %s.pdf', path_supplemental_figures, N), 6, 5)
  par(mfrow = c(2,2), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, xlab = '', yaxt = 'n', main = 'linear model')
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect z', 1, 2.5)
  abline(v = z_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
                                     summary(out[[i]]$linear_fit)$coefficients[3,2])), 
                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
  mtext('z', 2, 3, font = 2, las = 2, adj = 0)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = 'Bayesian linear model')
  abline(v = z_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_z), pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect z', 1, 2.5)

  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = '')
  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2], 
                                     summary(out[[i]]$linear_fit)$coefficients[2,2])), 
                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
  mtext('body', 2, 3, font = 2, las = 2, adj = 0)
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect body size', 1, 2.5)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = '')
  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), 
                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect body size', 1, 2.5)
  dev.off()
}
```

```{r plot response-20, echo=FALSE, out.width = "50%", fig.cap= "Model with sample size 20. Parameter estimates from the linear model and Bayesian linear model with brain size as response variable. Dashed grey line is the true value. Orange density plots are normal distributions based on the mean and SE from the linear model. Purple and green density plots are the posterior distributions from the Bayesian models."}
knitr::include_graphics(path_pdf_response_20)
```

```{r plot response-1000, echo=FALSE, out.width = "50%", fig.cap= "Model with sample size 1000. Parameter estimates from the linear model and Bayesian linear model with brain size as response variable. Dashed grey line is the true value. Orange density plots are normal distributions based on the mean and SE from the linear model. Purple and green density plots are the posterior distributions from the Bayesian models."}
knitr::include_graphics(path_pdf_response_1000)
```

### Case II: relative brain size as predictor variable

```{r predictor-sampsize, include=FALSE}
for(N in c(20, 1000)){

  # Settings
  set.seed(1)
  brain_effect = 1
  body_effect = 1
  model_predictor = cmdstan_model(path_predictor_stan)
  model_sem = cmdstan_model(path_sem_stan)
  if(N == 20) ylim = c(0, 5) else ylim = c(0, 15)

  # Simulate data
  out = lapply(1:20, function(i){
    body_size = rnorm(N, 0, 1)
    rel_brain_size = rnorm(N, 0, 1)
    brain_size = 1 * body_size + rel_brain_size
    mu_z = body_effect * body_size + brain_effect * rel_brain_size
    z = rnorm(N, mu_z, 1)

    # Run linear model
    linear_fit = lm(z ~ body_size + brain_size)

    # Run Bayesian linear model
    d = list(N = N, brain_size = brain_size, body_size = body_size, z = z)
    bay_linear_fit = model_predictor$sample(data = d, 
                                            seed = 1, 
                                            chains = 4, 
                                            parallel_chains = 4,
                                            refresh = 2000)
    post_linear_fit = bay_linear_fit$draws(format = 'draws_df')

    # Run Bayesian SEM
    bay_sem_fit = model_sem$sample(data = d, 
                                   seed = 1, 
                                   chains = 4, 
                                   parallel_chains = 4,
                                   refresh = 2000)
    post_sem_fit = bay_sem_fit$draws(format = 'draws_df')

    return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit, post_sem_fit = post_sem_fit))
  })

  # Plot the fit
  pdf(sprintf('%s/predictor %s.pdf', path_supplemental_figures, N), 9, 5)
  par(mfrow = c(2,3), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = 'linear model')
  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
                                     summary(out[[i]]$linear_fit)$coefficients[3,2])), 
                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect brain size', 1, 2.5)
  mtext('brain', 2, 4, font = 2, las = 2, adj = 0)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = 'Bayesian linear model')
  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_brain), 
                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect brain size', 1, 2.5)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = 'Bayesian SEM model')
  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_brain_z), 
                       pch = 16, cex = 3, col = alpha('#16A085', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect brain size', 1, 2.5)

  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = '')
  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2],
                                     summary(out[[i]]$linear_fit)$coefficients[2,2])), 
                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect body size', 1, 2.5)
  mtext('body', 2, 4, font = 2, las = 2, adj = 0)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = '')
  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), 
                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect body size', 1, 2.5)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = '')
  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_body_z), 
                       pch = 16, cex = 3, col = alpha('#16A085', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect body size', 1, 2.5)
  dev.off()

}
```

```{r plot predictor-20, echo=FALSE, out.width = "72%", fig.cap= "Model with sample size 20. Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable. Dashed grey line is the true value. Orange density plots are normal distributions based on the mean and SE from the linear model. Purple and green density plots are the posterior distributions from the Bayesian models."}
knitr::include_graphics(path_pdf_predictor_20)
```

```{r plot predictor-1000, echo=FALSE, out.width = "72%", fig.cap= "Model with sample size 1000. Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable. Dashed grey line is the true value. Orange density plots are normal distributions based on the mean and SE from the linear model. Purple and green density plots are the posterior distributions from the Bayesian models."}
knitr::include_graphics(path_pdf_predictor_1000)
```

## Priors

### Case I: relative brain size as response variable

```{r response-prior, include=FALSE}
for(prior_type in c('informative', 'vague')){

  # Settings
  set.seed(1)
  N = 20
  z_effect = 1
  body_effect = 1
  model_response = cmdstan_model(sprintf('%s/response_%s.stan', path_supplemental_models, prior_type))
  ylim = c(0, 5)

  # Simulate data 20 times
  out = lapply(1:20, function(i){
    body_size = rnorm(N, 1, 1)
    z = rnorm(N, 0, 1)
    mu_brain = body_effect * body_size + z_effect * z
    brain_size = rnorm(N, mu_brain, 1)

    # Run linear model
    linear_fit = lm(brain_size ~ body_size + z)

    # Run Bayesian linear model
    d = list(N = N, brain_size = brain_size, body_size = body_size, z = z) # save data as list
    bay_linear_fit = model_response$sample(data = d, # run model
                                           seed = 1, 
                                           chains = 4, 
                                           parallel_chains = 4, # runs on four threads
                                           refresh = 2000)
    post_linear_fit = bay_linear_fit$draws(format = 'draws_df') # draw posterior samples

    return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit))
  })

  # Plot the fit
  pdf(sprintf('%s/response %s.pdf', path_supplemental_figures, prior_type), 6, 5)
  par(mfrow = c(2,2), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, xlab = '', yaxt = 'n', main = 'linear model')
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect z', 1, 2.5)
  abline(v = z_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
                                     summary(out[[i]]$linear_fit)$coefficients[3,2])), 
                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
  mtext('z', 2, 3, font = 2, las = 2, adj = 0)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = 'Bayesian linear model')
  abline(v = z_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_z), pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect z', 1, 2.5)

  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = '')
  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2], 
                                     summary(out[[i]]$linear_fit)$coefficients[2,2])), 
                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
  mtext('body', 2, 3, font = 2, las = 2, adj = 0)
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect body size', 1, 2.5)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = '')
  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), 
                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect body size', 1, 2.5)
  dev.off()
}
```

```{r plot response-informative, echo=FALSE, out.width = "50%", fig.cap= "Model with informative priors. Parameter estimates from the linear model and Bayesian linear model with brain size as response variable. Dashed grey line is the true value. Orange density plots are normal distributions based on the mean and SE from the linear model. Purple and green density plots are the posterior distributions from the Bayesian models."}
knitr::include_graphics(path_pdf_response_informative)
```

```{r plot response-vague, echo=FALSE, out.width = "50%", fig.cap= "Model with vague priors. Parameter estimates from the linear model and Bayesian linear model with brain size as response variable. Dashed grey line is the true value. Orange density plots are normal distributions based on the mean and SE from the linear model. Purple and green density plots are the posterior distributions from the Bayesian models."}
knitr::include_graphics(path_pdf_response_vague)
```

### Case II: relative brain size as predictor variable

```{r predictor-prior, include=FALSE}
for(prior_type in c('informative', 'vague')){

  # Settings
  set.seed(1)
  N = 100
  brain_effect = 1
  body_effect = 1
  model_predictor = cmdstan_model(sprintf('%s/predictor_%s.stan', path_supplemental_models, prior_type))
  model_sem = cmdstan_model(sprintf('%s/sem_%s.stan', path_supplemental_models, prior_type))
  ylim = c(0, 5)

  # Simulate data
  out = lapply(1:20, function(i){
    body_size = rnorm(N, 0, 1)
    rel_brain_size = rnorm(N, 0, 1)
    brain_size = 1 * body_size + rel_brain_size
    mu_z = body_effect * body_size + brain_effect * rel_brain_size
    z = rnorm(N, mu_z, 1)

    # Run linear model
    linear_fit = lm(z ~ body_size + brain_size)

    # Run Bayesian linear model
    d = list(N = N, brain_size = brain_size, body_size = body_size, z = z)
    bay_linear_fit = model_predictor$sample(data = d, 
                                            seed = 1, 
                                            chains = 4, 
                                            parallel_chains = 4,
                                            refresh = 2000)
    post_linear_fit = bay_linear_fit$draws(format = 'draws_df')

    # Run Bayesian SEM
    bay_sem_fit = model_sem$sample(data = d, 
                                   seed = 1, 
                                   chains = 4, 
                                   parallel_chains = 4,
                                   refresh = 2000)
    post_sem_fit = bay_sem_fit$draws(format = 'draws_df')

    return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit, post_sem_fit = post_sem_fit))
  })

  # Plot the fit
  pdf(sprintf('%s/predictor %s.pdf', path_supplemental_figures, prior_type), 9, 5)
  par(mfrow = c(2,3), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = 'linear model')
  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
                                     summary(out[[i]]$linear_fit)$coefficients[3,2])), 
                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect brain size', 1, 2.5)
  mtext('brain', 2, 4, font = 2, las = 2, adj = 0)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = 'Bayesian linear model')
  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_brain), 
                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect brain size', 1, 2.5)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = 'Bayesian SEM model')
  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_brain_z), 
                       pch = 16, cex = 3, col = alpha('#16A085', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect brain size', 1, 2.5)

  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = '')
  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2],
                                     summary(out[[i]]$linear_fit)$coefficients[2,2])), 
                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect body size', 1, 2.5)
  mtext('body', 2, 4, font = 2, las = 2, adj = 0)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = '')
  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), 
                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect body size', 1, 2.5)
  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
       xlab = '', yaxt = 'n', main = '')
  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_body_z), 
                       pch = 16, cex = 3, col = alpha('#16A085', 0.5))
  axis(4)
  mtext('density', 4, 2.5)
  mtext('effect body size', 1, 2.5)
  dev.off()

}
```

```{r plot predictor-informative, echo=FALSE, out.width = "72%", fig.cap= "Model with informative priors. Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable. Dashed grey line is the true value. Orange density plots are normal distributions based on the mean and SE from the linear model. Purple and green density plots are the posterior distributions from the Bayesian models."}
knitr::include_graphics(path_pdf_predictor_informative)
```

```{r plot predictor-vague, echo=FALSE, out.width = "72%", fig.cap= "Model with vague priors. Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable. Dashed grey line is the true value. Orange density plots are normal distributions based on the mean and SE from the linear model. Purple and green density plots are the posterior distributions from the Bayesian models."}
knitr::include_graphics(path_pdf_predictor_vague)
```

### 2022-04-08 manuscript.Rmd
Starting to revise main text.

@@
 author: "Simeon Q. Smeele $^{1,2,3}$"
-date: "March 2022"
+date: "April 2022"
 output:
+  pdf_document: default
   word_document: default
-  pdf_document: default
 header-includes:
@@
 library(scales)
+library(daggity)
 
@@
 path_sem_stan = 'sem.stan'
-path_pdf_response = 'response.pdf' # pdfs
-path_pdf_predictor = 'predictor.pdf'
+path_pdf_response = 'figures/response.pdf' # pdfs
+path_pdf_predictor = 'figures/predictor.pdf'
 ```
@@
 
+```{r echo=FALSE}
+# Make DAG
+dag <- dagitty("dag {body_size -> relative_brain_size <- z}")
+plot( graphLayout( dag ) )
+```
+
 # Methods
@@
 Frequentist linear models were fitted with the *lm* function from base R [@R].
-Bayesian models were fitted using the *cmstanr* package [@cmdstanr], which runs the No U-turn Sampler in Stan [@gelman2015stan] with four chains and default sampler settings.
+Bayesian models were fitted using the *cmstanr* package [@cmdstanr], which runs the No U-turn Sampler in Stan [@gelman2015stan] with four chains and default sampler settings (1000 warmup and 1000 sampling iterations).
+Rhat, divergent transitions and effective sample size were monitored by the package and any issues were reported.
+
+To test if sample size had an effect on which model performed best, additional simulations were run and analysed for 20 and 1000 species (see supplemental results).
+To test if better or worse priors had an affect on which model performed best, and the performance of the best model, additional simulations were run with informative and vague priors (see supplemental materials).
 
@@
 Both the frequentist and Bayesian linear model estimated the effect of body mass to be essentially 0.
+
+Results were similar for simulations with smaller and larger sample sizes (see supplemental materials.
+Results were also similar for models with vague and informative prior (see supplemental materials).
+The informative priors constrained the Bayesian linear models estimate of the body size effect to positive values, but none of the posterior distributions included the true value of the body size effect.
+In other words, even if the model is run with the correct value as prior, it estimates the effect to be absent.
 

### 2022-04-16 manuscript.Rmd
Major update main text.

@@
 library(scales)
-library(daggity)
 
 # Paths
-path_response_stan = 'response.stan' # models
-path_predictor_stan = 'predictor.stan'
-path_sem_stan = 'sem.stan'
+path_response_stan = 'models/response.stan' # models
+path_predictor_stan = 'models/predictor.stan'
+path_sem_stan = 'models/sem.stan'
 path_pdf_response = 'figures/response.pdf' # pdfs
 path_pdf_predictor = 'figures/predictor.pdf'
+path_pdf_DAG = 'figures/DAG.pdf'
 ```
@@
 It is important to note that it is often the *relative* and not the *absolute* size of the brain that is of interest.
-When talking about relative brain size, it is always implicit that brain size is relative to body size or a specific brain region relative to the whole brain.
-This to control for allometric scaling.
+To control for allometric scaling (larger species have larger brains), brain size is measured relative to body size.
 There exists a large diversity of relative brain size measures [@healy2007critique], but some decades ago the use of residual brain size was proposed and quickly became the most popular [@jerison1973evolution].
@@
 Multiple challenges were recognised early on.
-Probably the most debated is the effect of phylogeny [@armstrong1983relative; @font2019rethinking; @burger2019allometry].
+Probably the most debated is the effect of phylogeny [@armstrong1983relative].
 When comparing across multiple species nested at different levels of taxonomy, how does one estimate the 'true' slope of the model?
-Several techniques have been proposed to mitigate this, but the debate is ongoing.
+Several techniques have been proposed to mitigate this, but the debate is ongoing [@font2019rethinking; @burger2019allometry].
 
 More recently another statistical caveat has been highlighted (for an overview see @freckleton2002misuse).
-The use of residuals is great for visualisation, but cannot be used in statistical models.
-The proposed solution is to include body size as a predictor variable in the final model, which is very similar to the use of residual brain size, if - and only if - relative brain size is the response variable.
+The use of residuals is great for visualisation, but should not be used as data.
+The proposed solution is to include body size in a multiple regression, which is very similar to the use of residual brain size, if - and only if - relative brain size is the response variable.
 This approach takes care of both the phylogenetic signal (since the model can include the phylogenetic variance-covariance matrix) and correct estimation of uncertainty.
-However, one major issue has been overlooked: if relative brain size is a predictor variable, including body size as a second predictor can lead to incorrect inference, because body size is not controlling for variation in brain size, since brain size is not the response variable.
-Brain size does, however, contain information about body size since body size is a cause of brain size.
-This will result in a reduction of the estimated effect of the body size.
-To properly include both relative brain size and body size one needs a system that contains regressions with brain size as response (of body size) and as predictor (of the third variable).
+However, one major issue has been overlooked: if relative brain size is a predictor variable, including body size as a second predictor leads to incorrect inference.
+This becomes clear when looking at the causal structure of such a case.
+The easiest way to decide which variables to include in an analysis is to draw a Directed Acyclic Graph (DAG).
+Here all variables are connected with arrows which indicate the direction of causality [@glymour2016causal].
+
+To point out the difference between the case with relative brain size as response vs predictor variable, I will present the underlying assumed causal structure presented in Figure 1.
+To avoid discussion about directionality I have simply named the third variable *z*.
+In the first case both body size and *z* cause brain size (see Figure 1a).
+To decide how to best estimate the direct effect of *z* on brain size, one needs to make sure that all *back-door paths* are closed.
+In other words, all arrows that points towards *z* need to be considered.
+But since there are no such arrows in this case, there is no need to include additional variables.
+However, body size explains most of the variation in brain size.
+Therefore it is still a good idea to include this variable to get a more precise estimate of the effect of *z* on brain size.
+Or in other words on relative brain size, since allometry is now accounted for.
+
+The second case is causally more opaque (see Figure 1b).
+In general the question of interest is how cognitive ability influences the third variable *z*.
+But since this cannot be measured reliably across species (therefore denoted by a U for unobserved), relative brain size is used as a proxy.
+This is where it becomes tricky.
+The assumption is that it is the *extra bit of brain* that is caused by the need for cognitive ability.
+And it is therefore *relative* brain size that needs to be included as proxy.
+The variable we initially measured is *absolute* brain size, and is a collider in this DAG.
+A collider is a variable that is caused by two or more other variables.
+When including a collider in the analysis it *opens up* a back-door path, in our case through body size.
+This is problematic, because the estimated effect of brain size on *z* now also contains some of the effect of body size on *z*.
+Therefore the estimated effect of body size on *z* will be biased.
+There is no way to correct for this using the variables in the DAG in a multiple regression.
+To include both relative brain size and body size one needs a system that contains regressions with brain size as response (of body size) and as predictor (of the third variable).
 
 A structural equation model is such a system.
-It contains regressions for each variable and allows brain size to be response and outcome variable simultaneously [@bowen2011structural].
-Structural equation models have been used before to include relative brain size in a comparative study in combination with phylogeny, measurement error and imputation [@smeele2021coevolution].
-The aim of this current paper is to show the estimation error in ordinary linear regressions using a simulation and propose a simple Bayesian structural equation model as a solution that can be easily adapted to most comparative studies.
+It contains regressions for each variable and allows brain size to be response and predictor variable simultaneously [@bowen2011structural].
+When fitted using a Bayesian approach, information flows in both directions, since the likelihood is computed for the whole system at each step.
+The aim of this paper is to show the estimation error in ordinary linear regressions using a simulation and propose a simple Bayesian structural equation model as a solution that can be easily adapted to most comparative studies.
+I also provide a version of the structural equation model with *ulam* from the *rethinking* package as front-end [@rethinking] so that models can be adapted within the R environment.
 
-```{r echo=FALSE}
-# Make DAG
-dag <- dagitty("dag {body_size -> relative_brain_size <- z}")
-plot( graphLayout( dag ) )
+```{r dag, echo=FALSE, out.width = "100%", fig.cap= "Directed Acyclic Graph of the two cases.
+Arrows indicatethe direction of causality.
+*z* stands for the third variable of interest.
+a) Case I: brain size is the reponse variable.
+b) Case II: relative brain size is a predictor variable.
+U stands for the unobserved variable cognitive ability."}
+knitr::include_graphics(path_pdf_DAG)
 ```
@@
 To show the difference between a case where relative brain size is the *response* variable and where it is a *predictor* variable I simulated two simple cases with three variables.
-To avoid discussion about directionality I have simply named the third variable *z*.
-To further simplify the simulation I have not attempted to simulate realistic values for body size and relative brain size, but just used values with mean = 0 and standard deviation = 1.
+To simplify the simulation I have not attempted to simulate realistic values for body size and relative brain size, but just used values with mean = 0 and standard deviation = 1.
 This allows me to draw general conclusions that are not sensitive to the scale of the variables.
@@
 To test if sample size had an effect on which model performed best, additional simulations were run and analysed for 20 and 1000 species (see supplemental results).
-To test if better or worse priors had an affect on which model performed best, and the performance of the best model, additional simulations were run with informative and vague priors (see supplemental materials).
+To test if more or less informative priors had an affect on which model performed best, and the performance of the best model, additional simulations were run with informative and vague priors (see supplemental materials).
 
@@
 z & \sim \text{normal}(\mu_z,\ \sigma_z) \\
-\mu_{z\text{[i]}} & = \alpha_z + \zeta_\text{body} * \text{body size}_\text{[i]} + \zeta_\text{brain} * (\text{brain size}_\text{[i]} - \mu_\text{brain[i]})\\
+\mu_{z\text{[i]}} & = \alpha_z + \gamma_\text{body} * \text{body size}_\text{[i]} + \gamma_\text{brain} * (\text{brain size}_\text{[i]} - \mu_\text{brain[i]})\\
 \\
 \alpha_\text{body},\ \alpha_\text{brain},\ \alpha_z  & \sim \text{normal}(0,\ 1)\\
-\beta_\text{body},\ \zeta_\text{body},\ \zeta_\text{brain} & \sim \text{normal}(0,\ 1)\\
+\beta_\text{body},\ \gamma_\text{body},\ \gamma_\text{brain} & \sim \text{normal}(0,\ 1)\\
 \sigma_\text{body},\ \sigma_\text{brain},\ \sigma_z & \sim \text{exponential}(1)\\
@@
 *z* is a function of body size and relative brain size (where relative brain size is the difference between the actual and expected brain size).
-Relative brain size in this model is very similar to residual brain size, but since it is computed at each iteration information flows in both directions and measurement error is correclty estimated.
+Relative brain size in this model is very similar to residual brain size, but since it is computed at each iteration information flows in both directions and measurement error is correctly estimated.
 The last three lines of the model are the priors for all parameters.
-Note that the priors for the slopes ($\beta$ and $\zeta$) are set to normal(0, 1), which regularises them slightly and assumes no effect of the predictors.
+Note that the priors for the slopes ($\beta$ and $\gamma$) are set to normal(0, 1), which regularises them slightly and assumes no effect of the predictors.
 For empirical studies theory might provide more informative priors, which would further increase the accuracy of the model.
@@
 
-For Case I, where relative brain size was the response variable, both models estimated the parameters very well (see Figure 1).
-For Case II, where relative brain size was a predictor variable, the effect of brain size was estimated well by all models, but the effect of body size was only estimated correctly by the structural equation model (see Figure 2).
+For Case I, where relative brain size was the response variable, both models estimated the parameters very well (see Figure 2).
+For Case II, where relative brain size was a predictor variable, the effect of brain size was estimated well by all models, but the effect of body size was only estimated correctly by the structural equation model (see Figure 3).
 Both the frequentist and Bayesian linear model estimated the effect of body mass to be essentially 0.
@@
 
-```{r plot response, echo=FALSE, out.width = "50%", fig.cap= "Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
+```{r plot response, echo=FALSE, out.width = "75%", fig.cap= "Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
 Dashed grey line is the true value.
@@
 
-  return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit, post_sem_fit = post_sem_fit))
+  return(list(linear_fit = linear_fit, 
+              post_linear_fit = post_linear_fit, 
+              post_sem_fit = post_sem_fit))
 })
@@
 
-```{r plot predictor, echo=FALSE, out.width = "72%", fig.cap= "Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable.
+```{r plot predictor, echo=FALSE, out.width = "100%", fig.cap= "Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable.
 Dashed grey line is the true value.
@@
 These studies also reported the direct effect of body size and sometimes drew conclusions based on the sign of this effect.
-Since all these studies used some version of a linear model (be it phylogenetic and/or Bayesian), they actually tested the effect of absolute brain size, as including body size only accounts for allometry in the response variable.
-The simulations in this paper showed that including body mass as additional predictor to control for allometric scaling of brain size works well if relative brain size is the response variable, but not if relative brain size is a predictor variable.
+Since all these studies used some version of a linear model (be it phylogenetic and/or Bayesian), they actually tested the effect of absolute brain size, since including body size only accounts for allometry in the response variable.
+The simulations in this paper showed that including body size as additional predictor to control for allometric scaling of brain size works well if relative brain size is the response variable, but not if relative brain size is a predictor variable.
 Perhaps counter-intuitively, the effect of brain size was still estimated correctly by all models.
@@
 One way to create some intuition about what is going on is that absolute brain size (which was the actual predictor variable included in the linear models) contains information about both body size and relative brain size.
-In a sense this variable controls for the body size effect already, leaving the direct effect of body size 0.
+In a sense this variable controls for the body size effect already, leaving the direct effect of body size zero.
 In a case where body size itself does not have an effect on *z*, it does not actually need to be included at all.
@@
 The use of these models is not limited to relative brain size, but can be used for any comparative study in which multiple causal paths are of interest.
+
+Structural equation models are similar to phylogenetic path analysis [@gonzalez2014introduction] and multivariate regressions [@izenman2013multivariate].
+However, both these techniques do not allow the inclusion of the difference between the observed and predicted brain size from one model as predictor for a second model.
+A potentially even more powerful approach was put forward by @smaers2009inferring and involves reconstructing the ancestral states of both body size and brain size.
+Recently this approach was used to test brain evolution in mammals and it was shown that relatively large brains can be achieved by highly divergent paths [@smaers2021evolution].
+The authors contest the notion that there is a universal scaling between body and brain size, and instead propose that relatively large brains can just as well be a result of selection towards smaller bodies.
+Despite this observation, there might still be value in the use of relative brain size for two reasons.
+First, the largest non-cognitive effects were due to a shift in the slope of the body to brain relation.
+Such a shift would be less likely to affect results when considering a lower order taxonomic group (e.g., only primates) and could be partially accounted for by including multiple slopes (e.g., one per family).
+Second, running a Bayesian multiregime OU modeling approach is not straightforward and becomes really difficult when many covariates are included.
+The best compromise is to first study the allometric patterns in the taxon of interest and then decide if residual brain size can be used as proxy for cognitive ability.
+
+For this paper I drew all variables from normal distributions with zero mean and standard deviation one.
+I furthermore simulated *z* as a function of relative brain size, rather than brain size and *z* as a function of the unobserved cognitive ability (as depicted in Figure 1b).
+I chose to do this to illustrate that even under these simple conditions a linear model cannot estimate the effect of body size correctly.
+In empirical studies, the exact causal structure will be different, but the general observation that one cannot control for body size in a linear model when relative brain size is theorised to predict *z* still stands.
+A regular confounder check as described by @laubach2021biologist should inform the design of the model.
+One example of relative brain size as predictor variable is the study of life history and innovativeness.
+@sol2016life suggested that relative brain size predicts both the innovation propensity and maximum life span of birds.
+They recognised the critique on using residuals, but decided to still use this approach, because they wanted to remove the allometric effects from brain size and not from innovation propensity.
+This was a valid argument since the orginal critique assumed a case where $x_1$ and $x_2$ are both causing $y$ [@freckleton2002misuse].
+When $x_1$ and $x_2$ are correlated, the use of a residual of $y \sim x_1$ creates a bias.
+In the case of life history and innovativeness the causal structure is different and this particular problem does not arise.
+However, in an empirical study measurement error, missing data and phylogenetic covariance still need to be accounted for and using residuals uncertainty does not propagate because models are run sequentially.
+
+Structural equation models allow for the inclusion of measurement error, imputation of missing data and phylogenetic coveriance [@smeele2022coevolution].
+Furthermore, the Bayesian implementation is useful in cases with low sample size.
+Prior information about parameters can be used to increase accuracy.
+E.g., as brain size can (on average) only increase with body size.
+The prior for this relationship should therfore be set to only include positive values.
+This creates a clear advantage over frequentist linear models, which allow any real number and can therefore produce scientifically nonsensical results.
+
+# Conclusion
+
+The aim of this paper was to show how a case where relative brain size is a response variable differs from a case where it is a predictor variable.
+From a simple simulation it becomes clear that one cannot include absolute brain size and body size as covariates if relative brain size is theorised to predict the third variable.
+Multiple other approaches can be used instead and the design of the model should be based on the assumed causal structure of the system.
+I provide a simple structural equation model, which works for the simulated data.
+Future studies should make sure all back-door paths are closed for their DAG and potentially include additional components to control for measurement error, missing data and phylogeny.
 
@@
 
-**Acknowledgements.** XXX
+**Acknowledgements.** I would like to thank all anonymous reviewers for valuable input. Their contribution helped to clarify the manuscript and gain new insights.
 

### 2022-04-16 run SEM with ulam as frontent.R
First version ulam model.

@@ -1,0 +1,44 @@
+# Settings
+library(rethinking)
+set.seed(1)
+N = 100
+brain_effect = 1
+body_effect = 1
+
+# Simulate data
+body_size = rnorm(N, 0, 1)
+rel_brain_size = rnorm(N, 0, 1)
+brain_size = 1 * body_size + rel_brain_size
+mu_z = body_effect * body_size + brain_effect * rel_brain_size
+z = rnorm(N, mu_z, 1)
+
+# Run Bayesian SEM using ulam from rethinking as front-end
+bay_sem_fit = ulam(alist(
+  
+  z ~ normal(mu_z, sigma_z), # z model
+  mu_z <- a_z + g_body * body_size + g_brain * (brain_size - mu_brain[i]),
+  
+  brain_size ~ normal(mu_brain, sigma_brain), # brain size model
+  mu_brain <- a_brain + b_body * body_size,
+  
+  body_size ~ normal(a_body, sigma_body), # body size model
+  
+  c(a_body, a_brain, a_z) ~ normal(0, 1),
+  c(b_body, g_body, g_brain) ~ normal(0, 1),
+  c(sigma_body, sigma_brain, sigma_z) ~ exponential(1)
+  
+), dat = list(body_size = body_size, brain_size = brain_size, z = z), # provide data as df or list
+chains = 4, cores = 4, iter = 2000, warmup = 500, cmdstan = TRUE) # settings for sampling
+
+print(precis(bay_sem_fit)) # print summary of results
+
+post_sem_fit = extract.samples(bay_sem_fit) # extract samples from the posterior
+
+# The post_sem_fit object has the same structure as the object in the main markdown and can also 
+# be plotted using the same code. 
+
+# The Stan-code can also be printed using the following line:
+stancode(bay_sem_fit)
+
+# For debugging the option sample = FALSE can be added to the ulam function so that Stan-code can be printed
+# even for models that fail. 

### 2022-04-18 manuscript.Rmd
Minor rewrites. Updating abstract to fit MEE format.

@@
 - \linenumbers
+- \usepackage{setspace}
+- \doublespacing
 editor_options:
@@
 
-**Subject areas:** evolution
-
-**Keywords:** relative brain size, comparative analysis, structural equation model, Bayesian
+**Keywords:** Bayesian statistics, comparative analysis, multiple regression, relative brain size, structural equation model
 
@@
 
+1.
 There is a long standing interest in the effect of relative brain size on other life history variables in a comparative context.
 Historically, residuals have been used to calculate these effects, but more recently it has been recognised that regression on residuals is not good practice.
-Instead, absolute brain size and body size are included in a single regression, with the idea that this controls for allometry.
-This practice leads to a second, potentially even greater estimation error when both relative brain size and body size are a cause of the third variable in question.
-In this study I use a simple simulation to show how the effect of body mass is masked if absolute brain size is included.
-I propose the use of structural equation models to simultaneously estimate relative brain size and its effect on the third variable, leaving the effect of body mass unconfounded.
+Instead, absolute brain size and body size are included in a multiple regression, with the idea that this controls for allometry.
+
+2.
+I use a simple simulation to illustrate how a case with brain size as response variable differs from a case with relative brain size as predictor variable.
+I use the simulated data to test which modelling approach can estimate the underlying causal effects for each case.
+
+3.
+The results show that a multiple regression models with both body size and another variable as predictor variable and brain size as response variable work well.
+However, if relative brain size is a predictor variable, a multiple regression fails to correctly estimate the effects.
+
+4. I propose the use of structural equation models to simultaneously estimate relative brain size and its effect on the third variable and discuss other potential methods.
 
@@
 When fitted using a Bayesian approach, information flows in both directions, since the likelihood is computed for the whole system at each step.
-The aim of this paper is to show the estimation error in ordinary linear regressions using a simulation and propose a simple Bayesian structural equation model as a solution that can be easily adapted to most comparative studies.
+The aim of this paper is to show the estimation error in multiple linear regressions using a simulation and propose a simple Bayesian structural equation model as a solution that can be easily adapted to most comparative studies.
 I also provide a version of the structural equation model with *ulam* from the *rethinking* package as front-end [@rethinking] so that models can be adapted within the R environment.
@@
 Arrows indicatethe direction of causality.
-*z* stands for the third variable of interest.
+z stands for the third variable of interest.
 a) Case I: brain size is the reponse variable.
-b) Case II: relative brain size is a predictor variable.
+b) Case II: relative brain size is a proxy predictor variable.
 U stands for the unobserved variable cognitive ability."}
@@
 I simulated 20 datasets per case and present parameter estimates for all datasets.
-Frequentist linear models were fitted with the *lm* function from base R [@R].
+Frequentist linear models were fitted with the *lm* function from R [@R].
 Bayesian models were fitted using the *cmstanr* package [@cmdstanr], which runs the No U-turn Sampler in Stan [@gelman2015stan] with four chains and default sampler settings (1000 warmup and 1000 sampling iterations).
@@
 In the second case both body size and relative brain size are a predictor of *z*.
-The interest of the study is to what extend relative brain size causes *z*.
+For simplicity I removed the unobserved variable from the simulation, and instead included a direct effect of relative brain size on z and absolute brain size.
+This would still be realistic when considering developmental time as a function of body size and relative brain size.
+The interest of the study is therefore to what extend relative brain size causes *z*.
 I simulated 20 data sets with 100 species with the following structure:
@@
 \text{relative brain size} & \sim \text{normal}(0,\ 1) \\
-\text{brain size} & = \text{relative brain size} + 1 * \text{body size} \\
+\text{brain size} & = 1 * \text{relative brain size} + 1 * \text{body size} \\
 z & \sim \text{normal}(\mu_z,\ 1) \\
@@
 Brain size is a function of body size.
-*z* is a function of body size and relative brain size (where relative brain size is the difference between the actual and expected brain size).
+*z* is a function of body size and relative brain size (where relative brain size is the difference between the actual and predicted brain size).
 Relative brain size in this model is very similar to residual brain size, but since it is computed at each iteration information flows in both directions and measurement error is correctly estimated.
@@
 
-Results were similar for simulations with smaller and larger sample sizes (see supplemental materials.
+Results were similar for simulations with smaller and larger sample sizes (see supplemental materials).
 Results were also similar for models with vague and informative prior (see supplemental materials).
@@
 # Plot the fit
-pdf(path_pdf_response, 6, 5)
+pdf(path_pdf_response, 7, 5.5)
 par(mfrow = c(2,2), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
@@
 
-```{r plot response, echo=FALSE, out.width = "75%", fig.cap= "Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
+```{r plot response, echo=FALSE, out.width = "72%", fig.cap= "Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
 Dashed grey line is the true value.
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple and green density plots are the posterior distributions from the Bayesian models."}
+Purple density plots are the posterior distributions from the Bayesian model."}
 knitr::include_graphics(path_pdf_response)
@@
 # Plot the fit
-pdf(path_pdf_predictor, 9, 5)
+pdf(path_pdf_predictor, 9, 4.8)
 par(mfrow = c(2,3), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
@@
 
-```{r plot predictor, echo=FALSE, out.width = "100%", fig.cap= "Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable.
+```{r plot predictor, echo=FALSE, out.width = "100%", fig.cap= "Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
 Dashed grey line is the true value.
@@
 Using absolute brain size would of course lead to a less precise estimate of the effect of brain size, so using relative brain size from a structural equation model would still be preferable.
-The use of these models is not limited to relative brain size, but can be used for any comparative study in which multiple causal paths are of interest.
 
@@
 For this paper I drew all variables from normal distributions with zero mean and standard deviation one.
-I furthermore simulated *z* as a function of relative brain size, rather than brain size and *z* as a function of the unobserved cognitive ability (as depicted in Figure 1b).
-I chose to do this to illustrate that even under these simple conditions a linear model cannot estimate the effect of body size correctly.
-In empirical studies, the exact causal structure will be different, but the general observation that one cannot control for body size in a linear model when relative brain size is theorised to predict *z* still stands.
+I furthermore simulated *z* as a function of relative brain size, rather than simulating both *z* and brain size as a function of the unobserved cognitive ability (as depicted in Figure 1b).
+I chose to do this to illustrate that even under these simple conditions a multiple linear regression cannot estimate the effect of body size correctly.
+In empirical studies the exact causal structure will be different, but the general observation that one cannot control for allometry in a multiple linear regression when brain size is not the response variable still stands.
 A regular confounder check as described by @laubach2021biologist should inform the design of the model.
@@
 @sol2016life suggested that relative brain size predicts both the innovation propensity and maximum life span of birds.
-They recognised the critique on using residuals, but decided to still use this approach, because they wanted to remove the allometric effects from brain size and not from innovation propensity.
-This was a valid argument since the orginal critique assumed a case where $x_1$ and $x_2$ are both causing $y$ [@freckleton2002misuse].
+They recognised the critique on using residuals, but still used this approach, because they wanted to remove the allometric effects from brain size and not from innovation propensity.
+This was a valid argument since the original critique assumed a case where $x_1$ and $x_2$ are both causing $y$ [@freckleton2002misuse].
 When $x_1$ and $x_2$ are correlated, the use of a residual of $y \sim x_1$ creates a bias.
 In the case of life history and innovativeness the causal structure is different and this particular problem does not arise.
-However, in an empirical study measurement error, missing data and phylogenetic covariance still need to be accounted for and using residuals uncertainty does not propagate because models are run sequentially.
+However, in an empirical study measurement error, missing data and phylogenetic covariance still need to be accounted for.
+When using residuals, uncertainty does not propagate because models are run sequentially.
 
@@
 Prior information about parameters can be used to increase accuracy.
-E.g., as brain size can (on average) only increase with body size.
-The prior for this relationship should therfore be set to only include positive values.
+E.g., on average brain size can only increase with body size.
+The prior for this relationship should therefore be set to only include positive values.
 This creates a clear advantage over frequentist linear models, which allow any real number and can therefore produce scientifically nonsensical results.
@@
 Future studies should make sure all back-door paths are closed for their DAG and potentially include additional components to control for measurement error, missing data and phylogeny.
+The use of these models is not limited to relative brain size, but can be used for any comparative study in which multiple causal paths are of interest.
 

### 2022-04-18 manuscript.Rmd
Minor updates.

@@
 author: "Simeon Q. Smeele $^{1,2,3}$"
-date: "April 2022"
 output:
@@
 $^3$Department of Biology, University of Konstanz, Konstanz, Germany
-
-**ORCID:** 0000-0003-1001-6615
 
@@
 
-Results were similar for simulations with smaller and larger sample sizes (see supplemental materials).
-Results were also similar for models with vague and informative prior (see supplemental materials).
+Results were similar for simulations with smaller and larger sample sizes (see supplemental materials, Figure S1-4).
+Results were also similar for models with vague and informative prior (see supplemental materials, Figure S5-8).
 The informative priors constrained the Bayesian linear models estimate of the body size effect to positive values, but none of the posterior distributions included the true value of the body size effect.
@@
 
+**Acknowledgements.** I would like to thank all anonymous reviewers for valuable input. Their contribution helped to clarify the manuscript and gain new insights.
+
 **Data accessibility.** All data is generated in the simulation. Code is publicly available at: https://github.com/simeonqs/On_the_use_of_relative_brain_size.
@@
 
-**Acknowledgements.** I would like to thank all anonymous reviewers for valuable input. Their contribution helped to clarify the manuscript and gain new insights.
+**ORCID:** 0000-0003-1001-6615
 

### 2022-04-18 supplemental materials.Rmd
Updating all pdfs.

@@
 - \linenumbers
+- \renewcommand{\figurename}{Figure S}
+- \makeatletter
+- \def\fnum@figure{\figurename\thefigure}
+- \makeatother
+- \usepackage{setspace}
+- \doublespacing
 editor_options:
@@
 
+\pagebreak
+
 # Results
-
-## Sample size
-
-### Case I: relative brain size as response variable
 
@@
 
-```{r plot response-20, echo=FALSE, out.width = "50%", fig.cap= "Model with sample size 20.
+```{r plot response-20, echo=FALSE, out.width = "45%", fig.cap= "Model with sample size 20.
 Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
@@
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple and green density plots are the posterior distributions from the Bayesian models."}
+Purple density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_response_20)
@@
 
-```{r plot response-1000, echo=FALSE, out.width = "50%", fig.cap= "Model with sample size 1000.
+```{r plot response-1000, echo=FALSE, out.width = "45%", fig.cap= "Model with sample size 1000.
 Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
@@
 ```
-
-### Case II: relative brain size as predictor variable
 
@@
 
-```{r plot predictor-20, echo=FALSE, out.width = "72%", fig.cap= "Model with sample size 20.
-Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable.
+```{r plot predictor-20, echo=FALSE, out.width = "65%", fig.cap= "Model with sample size 20.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model  with relative brain size as predictor variable.
 Dashed grey line is the true value.
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple and green density plots are the posterior distributions from the Bayesian models."}
+Purple density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_predictor_20)
@@
 
-```{r plot predictor-1000, echo=FALSE, out.width = "72%", fig.cap= "Model with sample size 1000.
-Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable.
+```{r plot predictor-1000, echo=FALSE, out.width = "65%", fig.cap= "Model with sample size 1000.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model  with relative brain size as predictor variable.
 Dashed grey line is the true value.
@@
 ```
-
-## Priors
-
-### Case I: relative brain size as response variable
 
@@
 
-```{r plot response-informative, echo=FALSE, out.width = "50%", fig.cap= "Model with informative priors.
+```{r plot response-informative, echo=FALSE, out.width = "45%", fig.cap= "Model with informative priors.
 Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
@@
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple and green density plots are the posterior distributions from the Bayesian models."}
+Purple density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_response_informative)
@@
 
-```{r plot response-vague, echo=FALSE, out.width = "50%", fig.cap= "Model with vague priors.
+```{r plot response-vague, echo=FALSE, out.width = "45%", fig.cap= "Model with vague priors.
 Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
@@
 
-```{r plot predictor-informative, echo=FALSE, out.width = "72%", fig.cap= "Model with informative priors.
-Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable.
+```{r plot predictor-informative, echo=FALSE, out.width = "65%", fig.cap= "Model with informative priors.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model  with relative brain size as predictor variable.
 Dashed grey line is the true value.
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple and green density plots are the posterior distributions from the Bayesian models."}
+Purple density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_predictor_informative)
@@
 
-```{r plot predictor-vague, echo=FALSE, out.width = "72%", fig.cap= "Model with vague priors.
-Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model (SEM) with relative brain size as predictor variable.
+```{r plot predictor-vague, echo=FALSE, out.width = "65%", fig.cap= "Model with vague priors.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model  with relative brain size as predictor variable.
 Dashed grey line is the true value.

### 2022-07-09 manuscript.Rmd
Major update to main text after comments from Ecology and Evolution reviewers.

@@
   pdf_document: default
-  word_document: default
 header-includes:
@@
 library(scales)
+library(knitr)
 
@@
 path_pdf_DAG = 'figures/DAG.pdf'
+
+# Table captions
+tn = local({
+  i = 0
+  function(x) {
+    i <<- i + 1
+    paste('\n\n:Table ', i, ': ', x, sep = '')
+    # The : before Table tells pandoc to wrap your caption in <caption></caption>
+  }
+})
+knit_hooks$set(tab.cap = function(before, options, envir) {
+  if(!before)
+    tn(options$tab.cap)
+})
+default_output_hook = knit_hooks$get("output")
+knit_hooks$set(output = function(x, options) {
+  if (is.null(options$tab.cap) == F)  
+    x
+  else
+    default_output_hook(x,options)
+})
 ```
@@
 
-**Author for correspondence:** Simeon Q. Smeele, ssmeele@ab.mpg.de
+**Author for correspondence:** Simeon Q. Smeele, Max Planck Institute of Animal Behavior, Am Obstberg 1, 78315 Radolfzell am Bodensee, ssmeele@ab.mpg.de
 
@@
 3.
-The results show that a multiple regression models with both body size and another variable as predictor variable and brain size as response variable work well.
-However, if relative brain size is a predictor variable, a multiple regression fails to correctly estimate the effects.
+The results show that a multiple regression model with both body size and another variable as predictor variable and brain size as response variable work well.
+However, if relative brain size is a predictor variable, a multiple regression fails to correctly estimate the effect of body size.
 
@@
 
-Brain size is often theorised to be important in the life history of species and has been linked to everything from innovation rates [@lefebvre2004brains], to sociality [@dunbar2007evolution] and longevity [@minias2017longevity].
+Brain size is often theorised to be important in the life history of species and has been linked to many traits such as innovation rates [@lefebvre2004brains], sociality [@dunbar2007evolution] and longevity [@minias2017longevity].
 It is important to note that it is often the *relative* and not the *absolute* size of the brain that is of interest.
@@
 There exists a large diversity of relative brain size measures [@healy2007critique], but some decades ago the use of residual brain size was proposed and quickly became the most popular [@jerison1973evolution].
-This measure is simply the residual from a model that has log body size as predictor and log brain size as response variable.
-It is an intuitively attractive approach, since it seems to get at *that extra bit of brain* a species has.
+This measure is the residual from a model that has log body size as predictor and log brain size as response variable.
+It is an intuitively attractive approach, since it measures whether the brain is smaller or larger than expected from allometry alone.
 Multiple challenges were recognised early on.
@@
 More recently another statistical caveat has been highlighted (for an overview see @freckleton2002misuse).
-The use of residuals is great for visualisation, but should not be used as data.
+The use of residuals leads to biased estimates if correlation exists between the predictor variables, as is often the case in the comparative literature.
 The proposed solution is to include body size in a multiple regression, which is very similar to the use of residual brain size, if - and only if - relative brain size is the response variable.
@@
 This becomes clear when looking at the causal structure of such a case.
-The easiest way to decide which variables to include in an analysis is to draw a Directed Acyclic Graph (DAG).
-Here all variables are connected with arrows which indicate the direction of causality [@glymour2016causal].
+The best way to decide which variables to include in an analysis is the use of causal graphs [@glymour2016causal, @wright1934method].
+@laubach2021biologist presented a simple guide to draw Directed Acyclic Graphs (DAG), in which all variables are connected with arrows that indicate the direction of causality.
 
@@
 When fitted using a Bayesian approach, information flows in both directions, since the likelihood is computed for the whole system at each step.
-The aim of this paper is to show the estimation error in multiple linear regressions using a simulation and propose a simple Bayesian structural equation model as a solution that can be easily adapted to most comparative studies.
+The aim of this paper is to show the estimation bias in multiple linear regressions using a simulation and propose a simple Bayesian structural equation model as a solution that can be easily adapted to most comparative studies.
 I also provide a version of the structural equation model with *ulam* from the *rethinking* package as front-end [@rethinking] so that models can be adapted within the R environment.
@@
 z stands for the third variable of interest.
-a) Case I: brain size is the reponse variable.
+a) Case I: brain size is the response variable.
 b) Case II: relative brain size is a proxy predictor variable.
@@
 This allows me to draw general conclusions that are not sensitive to the scale of the variables.
+It should also be noted that body and brain size are normally measured on the natural scale, but log-transformed before analysis.
+The asumption is that causal effects are linear on the logarithmic scale.
+I have therefore simulated both body and brain size to be on the logarithmic scale.
 
@@
 Frequentist linear models were fitted with the *lm* function from R [@R].
-Bayesian models were fitted using the *cmstanr* package [@cmdstanr], which runs the No U-turn Sampler in Stan [@gelman2015stan] with four chains and default sampler settings (1000 warmup and 1000 sampling iterations).
-Rhat, divergent transitions and effective sample size were monitored by the package and any issues were reported.
+Bayesian models were fitted using the *cmdstanr* package [@cmdstanr], which runs the No U-turn Sampler in Stan [@gelman2015stan] with four chains and default sampler settings (1000 warmup and 1000 sampling iterations).
+Rhat, divergent transitions and effective sample size were monitored by the package and if issues arose these were reported in the results.
 
 To test if sample size had an effect on which model performed best, additional simulations were run and analysed for 20 and 1000 species (see supplemental results).
-To test if more or less informative priors had an affect on which model performed best, and the performance of the best model, additional simulations were run with informative and vague priors (see supplemental materials).
+To test if more or less informative priors had an effect on which model performed best, and how often the best model included the true effect in the credible interval, additional simulations were run with informative and vague priors (see supplemental materials).
 
@@
 In the first case absolute brain size is caused by body size and *z*.
-The interest of the study is to what extent *z* causes additional increase in brain size.
-In other words, to what extend *z* correlates with relative brain size.
+The interest of the study is to what extent *z* causes additional variation in brain size (when the effect of body size is accounted for).
+In other words, to what extent *z* correlates with relative brain size.
 I simulated 20 data sets with 100 species with the following structure:
@@
 
-I analysed the resulting data with a frequentist linear model and with the Bayesian equivalent. Then I plotted the estimated effect of body size and *z* to show how well parameters were retrieved.
+I analysed the resulting data with a frequentist linear model and with the Bayesian equivalent.
+Then I plotted the estimated coefficient of body size and *z* to show how well parameters were retrieved.
+Additionally the bias (difference between estimate and simulated value) and coverage (proportion of times the true value was within 2SE distance from the estimate) are reported.
 
@@
 For simplicity I removed the unobserved variable from the simulation, and instead included a direct effect of relative brain size on z and absolute brain size.
-This would still be realistic when considering developmental time as a function of body size and relative brain size.
-The interest of the study is therefore to what extend relative brain size causes *z*.
+In this way absolute brain size is the sum of the brain tissue needed for enervating the body and the additional brain tissue, which is ultimately caused by other variables.
+Such a situation is still realistic when considering developmental time as a function of body size and relative brain size.
+The interest of the study is therefore to what extent relative brain size causes *z*.
 I simulated 20 data sets with 100 species with the following structure:
@@
 The last three lines of the model are the priors for all parameters.
-Note that the priors for the slopes ($\beta$ and $\gamma$) are set to normal(0, 1), which regularises them slightly and assumes no effect of the predictors.
+Note that the priors for the slopes ($\beta$ and $\gamma$) are set to normal(0, 1), which regularises them slightly and are centred no effect of the predictors.
 For empirical studies theory might provide more informative priors, which would further increase the accuracy of the model.
+
+To test how models performed in case of unqual magnitute of the effect of body size and brain size on *z* (case II), additional simulation were run (see supplemental materials) with either strong body size effect ($\beta_\text{body} = 2, \beta_\text{brain} = 0.5$) or strong brain size effect ($\beta_\text{body} = 0.5, \beta_\text{brain} = 2$).
+
+To visualise the results I plotted all parameter estimates of the main model (for *z*) and reported the bias (difference between estimate and simulated value) and coverage (proportion of times the true value was within 2SE distance from the estimate).
 
@@
 
-For Case I, where relative brain size was the response variable, both models estimated the parameters very well (see Figure 2).
-For Case II, where relative brain size was a predictor variable, the effect of brain size was estimated well by all models, but the effect of body size was only estimated correctly by the structural equation model (see Figure 3).
+For Case I, where relative brain size was the response variable, both models estimated the parameters very well (see Figure 2, Table 1).
+For Case II, where relative brain size was a predictor variable, the effect of brain size was estimated well by all models, but the effect of body size was only estimated correctly by the structural equation model (see Figure 3, Table 1).
 Both the frequentist and Bayesian linear model estimated the effect of body mass to be essentially 0.
@@
 The informative priors constrained the Bayesian linear models estimate of the body size effect to positive values, but none of the posterior distributions included the true value of the body size effect.
-In other words, even if the model is run with the correct value as prior, it estimates the effect to be absent.
+In other words, even if the model is run with the prior centred tightly around the correct value, it estimates the effect to be absent.
+Also when varying the magnitude of the effects of body size and brain size, the structural equation model was the only model to correctly estimate the effect of body size (see S9-10).
 
@@
 dev.off()
+
+# Calculate bias 
+bias_response_body_linear = sapply(out, function(x) body_effect - x$linear_fit$coefficients[2])
+bias_response_z_linear = sapply(out, function(x) z_effect - x$linear_fit$coefficients[3])
+bias_response_body_linear_bayes = sapply(out, function(x) body_effect - mean(x$post_linear_fit$b_body))
+bias_response_z_linear_bayes = sapply(out, function(x) z_effect - mean(x$post_linear_fit$b_z))
+max_abs_bias = max(abs(c(bias_response_body_linear, 
+                         bias_response_z_linear,
+                         bias_response_body_linear_bayes,
+                         bias_response_z_linear_bayes)))
+
+# Calculate coverage
+calc.cov = function(true_value, mean, SE){
+  if(abs(mean - true_value) < 2*SE) return(T) else return(F)
+}
+coverage_response_body_linear = sapply(out, function(x) 
+  calc.cov(body_effect, x$linear_fit$coefficients[2], summary(out[[i]]$linear_fit)$coefficients[2,2]))
+coverage_response_z_linear = sapply(out, function(x) 
+  calc.cov(z_effect, x$linear_fit$coefficients[3], summary(out[[i]]$linear_fit)$coefficients[3,2]))
+coverage_response_body_linear_bayes = sapply(out, function(x) 
+  calc.cov(body_effect, mean(x$post_linear_fit$b_body), sd(x$post_linear_fit$b_body)))
+coverage_response_z_linear_bayes = sapply(out, function(x) 
+  calc.cov(z_effect, mean(x$post_linear_fit$b_z), sd(x$post_linear_fit$b_z)))
+
+# Make table
+table_response = data.frame(
+  case = rep('response', 4),
+  parameter = c('body', 'body', 'z', 'z'),
+  model = c('linear', 'Bayesian linear', 'linear', 'Bayesian linear'),
+  mean_bias = c(mean(bias_response_body_linear),
+                mean(bias_response_body_linear_bayes),
+                mean(bias_response_z_linear),
+                mean(bias_response_z_linear_bayes)),
+  sd_bias = c(sd(bias_response_body_linear),
+              sd(bias_response_body_linear_bayes),
+              sd(bias_response_z_linear),
+              sd(bias_response_z_linear_bayes)),
+  coverage = c(sum(coverage_response_body_linear)/length(coverage_response_body_linear),
+               sum(coverage_response_body_linear_bayes)/length(coverage_response_body_linear_bayes),
+               sum(coverage_response_z_linear)/length(coverage_response_z_linear),
+               sum(coverage_response_z_linear_bayes)/length(coverage_response_z_linear_bayes)))
+table_response[,4:6] = round(table_response[,4:6], 2)
 ```
@@
               post_linear_fit = post_linear_fit, 
-              post_sem_fit = post_sem_fit))
+              post_sem_fit = post_sem_fit,
+              body_size = body_size,
+              brain_size = brain_size))
 })
@@
 mtext('density', 4, 2.5)
-mtext('effect brain size', 1, 2.5)
+mtext('effect relative brain size', 1, 2.5)
 
@@
 dev.off()
+
+# Calculate bias 
+bias_predictor_body_linear = sapply(out, function(x) body_effect - x$linear_fit$coefficients[2])
+bias_predictor_brain_linear = sapply(out, function(x) brain_effect - x$linear_fit$coefficients[3])
+bias_predictor_body_linear_bayes = sapply(out, function(x) body_effect - mean(x$post_linear_fit$b_body))
+bias_predictor_brain_linear_bayes = sapply(out, function(x) brain_effect - mean(x$post_linear_fit$b_brain))
+bias_predictor_body_sem = sapply(out, function(x) body_effect - mean(x$post_sem_fit$b_body_z))
+bias_predictor_brain_sem = sapply(out, function(x) brain_effect - mean(x$post_sem_fit$b_brain_z))
+
+# Calculate coverage
+calc.cov = function(true_value, mean, SE){
+  if(abs(mean - true_value) < 2*SE) return(T) else return(F)
+}
+coverage_predictor_body_linear = sapply(out, function(x) 
+  calc.cov(body_effect, x$linear_fit$coefficients[2], summary(out[[i]]$linear_fit)$coefficients[2,2]))
+coverage_predictor_brain_linear = sapply(out, function(x) 
+  calc.cov(brain_effect, x$linear_fit$coefficients[3], summary(out[[i]]$linear_fit)$coefficients[3,2]))
+coverage_predictor_body_linear_bayes = sapply(out, function(x) 
+  calc.cov(body_effect, mean(x$post_linear_fit$b_body), sd(x$post_linear_fit$b_body)))
+coverage_predictor_brain_linear_bayes = sapply(out, function(x) 
+  calc.cov(brain_effect, mean(x$post_linear_fit$b_brain), sd(x$post_linear_fit$b_brain)))
+coverage_predictor_body_sem = sapply(out, function(x) 
+  calc.cov(body_effect, mean(x$post_sem_fit$b_body_z), sd(x$post_sem_fit$b_body_z)))
+coverage_predictor_brain_sem = sapply(out, function(x) 
+  calc.cov(brain_effect, mean(x$post_sem_fit$b_brain_z), sd(x$post_sem_fit$b_brain_z)))
+
+# Make table
+table_predictor = data.frame(
+  case = rep('predictor', 6),
+  parameter = c('body', 'body', 'body', 'brain', 'brain', 'brain'),
+  model = c('linear', 'Bayesian linear', 'SEM', 'linear', 'Bayesian linear',
+            'SEM'),
+  mean_bias = c(mean(bias_predictor_body_linear),
+                mean(bias_predictor_body_linear_bayes),
+                mean(bias_predictor_body_sem),
+                mean(bias_predictor_brain_linear),
+                mean(bias_predictor_brain_linear_bayes),
+                mean(bias_predictor_brain_sem)),
+  sd_bias = c(sd(bias_predictor_body_linear),
+              sd(bias_predictor_body_linear_bayes),
+              sd(bias_predictor_body_sem),
+              sd(bias_predictor_brain_linear),
+              sd(bias_predictor_brain_linear_bayes),
+              sd(bias_predictor_brain_sem)),
+  coverage = c(sum(coverage_predictor_body_linear)/length(coverage_predictor_body_linear),
+               sum(coverage_predictor_body_linear_bayes)/length(coverage_predictor_body_linear_bayes),
+               sum(coverage_predictor_body_sem)/ length(coverage_predictor_body_sem),
+               sum(coverage_predictor_brain_linear)/length(coverage_predictor_brain_linear),
+               sum(coverage_predictor_brain_linear_bayes)/length(coverage_predictor_brain_linear_bayes),
+               sum(coverage_predictor_brain_sem)/length(coverage_predictor_brain_sem)))
+table_predictor[,4:6] = round(table_predictor[,4:6], 2)
+
+# Calculate average correlation between body size and absolute brain size
+correlation = sapply(out, function(x) cor(x$body_size, x$brain_size))
+mean_correlation = mean(correlation)
 ```
@@
 
+```{r table, echo=FALSE}
+table_combined = rbind(table_response, table_predictor)
+knitr::kable(table_combined, format = 'markdown', caption = 'Bias and coverage of all models.
+Mean and standard deviation bias is calculated from the distance between the point estimate and simulated value.
+Coverage is the proportion of times that the simulated value was withing the point estimate +- 2SE.')
+```
+
 # Discussion
 
-Several recent studies have claimed to study the effect of relative brain size by including both absolute brain size and body size as predictors [@street2017coevolution; @gonzalez2010large; @isler2009why].
+Several recent studies have claimed to study the effect of relative brain size by including both absolute brain size and body size as predictors [@street2017coevolution; @gonzalez2010large; @isler2009why, @maklakov2011brains, @samia2015brain].
 These studies also reported the direct effect of body size and sometimes drew conclusions based on the sign of this effect.
@@
 It was the body size effect that was biased in the linear models.
-The use of such models would lead to the incorrect inference that there was no effect of body size at all.
+Simulations with a strong body size effect still found a positive coefficient for this effect, but were biased towards lower values (see Figure S9).
+In simulations with a strong brain size effect the body size coefficients were even negative (see Figure S10).
 
 One way to create some intuition about what is going on is that absolute brain size (which was the actual predictor variable included in the linear models) contains information about both body size and relative brain size.
-In a sense this variable controls for the body size effect already, leaving the direct effect of body size zero.
-In a case where body size itself does not have an effect on *z*, it does not actually need to be included at all.
-Using absolute brain size would be fine.
+In a sense this variable controls for part of the body size effect already.
+In other words, there are two paths from relative brain size (or the unobserved cognitive ability from the DAG) to *z*: one path is direct and one path is a back-door path through the allometric effect of body size.
+In another case, where body size itself does not have an effect on *z*, it does not actually need to be included at all.
+In this simple example, using absolute brain size would be fine.
 The variation in brain size due to allometric scaling would just create noise.
-Using absolute brain size would of course lead to a less precise estimate of the effect of brain size, so using relative brain size from a structural equation model would still be preferable.
+However, in empirical studies the causal structure is often more complex, and researchers should always check if using absolute brain size leads to bias.
+Using absolute brain size would also lead to a less precise estimate of the effect of brain size, so using relative brain size from a structural equation model would still be preferable.
 
@@
 Such a shift would be less likely to affect results when considering a lower order taxonomic group (e.g., only primates) and could be partially accounted for by including multiple slopes (e.g., one per family).
-Second, running a Bayesian multiregime OU modeling approach is not straightforward and becomes really difficult when many covariates are included.
-The best compromise is to first study the allometric patterns in the taxon of interest and then decide if residual brain size can be used as proxy for cognitive ability.
+Second, running a Bayesian multi-regime OU modelling approach is not straightforward and becomes really difficult when many covariates are included, since no step-by-step guide or R package exists.
+A compromise is to first study the allometric patterns in the taxon of interest and then decide if relative brain size can be used as proxy for cognitive ability.
+
+The use of an artificial variable *z* might be confusing, so it might be informative to give examples for both cases and explain the evolutionary causal structure.
+The social brain hypothesis is possible the most hotly debated version of case I, with relative brain size as response variable.
+Here brain size is a function of body size and some measure of sociality (*z*).
+The idea is that social species experience evolutionary pressure to increase brain size to deal with complex social situations.
+First brought up by @jolly1966lemur, this hypotheses has since received much attention both with residual brain size and absolute brain size as response variable (for a review see @acedo2016critical).
+The second case, where relative brain size is a predictor variable, has received less attention, but several studies have attempted to show that relative brain size can cause longevity, with relatively larger brained species living longer [@street2017coevolution, @yu2018large, @decasien2018encephalization, @jimenez2020long].
+The understanding is that cognitive ability causes both increased brain size and the ability to survive longer.
+This might be unintuitive, since mechanistically cognition is caused by the brain, but in an evolutionary context it makes sense, since species that require more cognitive flexibility experience evolutionary pressure to increase brain size.
 
@@
 I furthermore simulated *z* as a function of relative brain size, rather than simulating both *z* and brain size as a function of the unobserved cognitive ability (as depicted in Figure 1b).
-I chose to do this to illustrate that even under these simple conditions a multiple linear regression cannot estimate the effect of body size correctly.
+I chose to do this to illustrate that even under simple conditions a multiple linear regression cannot estimate the effect of body size correctly.
 In empirical studies the exact causal structure will be different, but the general observation that one cannot control for allometry in a multiple linear regression when brain size is not the response variable still stands.
@@
 However, in an empirical study measurement error, missing data and phylogenetic covariance still need to be accounted for.
-When using residuals, uncertainty does not propagate because models are run sequentially.
+With the use of a structural equation model one obtains direct estimates of relative brain size, which includes uncertainty stemming from measurement uncertainty on body size and on brain size.
+This can also be plotted directly, allowing researchers to visualise how certain the correlation of relative brain size with the third variable is.
 
-Structural equation models allow for the inclusion of measurement error, imputation of missing data and phylogenetic coveriance [@smeele2022coevolution].
+Another problem with the use of residuals was brought up by @rogell2020controlling.
+They show that many studies have oposite signs between the model with absolute brain size and the model with residual brain size.
+They argue that coefficients from linear models are difficult to interpret in a biologically meaningfull way, when body size and brain size are strongly correlated.
+In the current study, the average correlation between body size and brain size was `r mean_correlation`, which is not as strong as in some emperical studies, but still cleary problematic for the linear models.
+The SEM estimates did, however, not suffer under these conditions.
+
+Structural equation models allow for the inclusion of measurement error, imputation of missing data and phylogenetic coveriance.
+@smeele2022coevolution published their R script for such an applied structural equation model, which can be modified to fit most comparative questions.
 Furthermore, the Bayesian implementation is useful in cases with low sample size.
 Prior information about parameters can be used to increase accuracy.
-E.g., on average brain size can only increase with body size.
+E.g., it is highly unlikely that an increase in body size would directly lead to a decrease in body size.
 The prior for this relationship should therefore be set to only include positive values.

### 2022-07-09 supplemental materials.Rmd
Major update to main text after comments from Ecology and Evolution reviewers.

@@
 author: "Simeon Q. Smeele $^{1,2,3}$"
-date: "April 2022"
+date: "July 2022"
 output:
@@
 path_pdf_predictor_vague = 'supplemental figures/predictor vague.pdf'
+path_pdf_predictor_strong_body = 'supplemental figures/predictor strong body.pdf'
+path_pdf_predictor_strong_brain = 'supplemental figures/predictor strong brain.pdf'
 ```
@@
 To further test the robustness of the models I analysed the data with vague priors, settings intercept and slope to normal(0, 10) and standard variation parameters to exponential(0.1) (note that the mean for the exponential distribution is the inverse of the rate, such that the mean standard deviation here is 10).
+
+## Effect magnitude
+
+The main text assumes equal magnitude for the body size and brain size effect.
+To test how the models performed when the body and brain effects on z are unequal in magnitude, I ran two additional simulations with either strong body size effect ($\beta_\text{body} = 2, \beta_\text{brain} = 0.5$) or strong brain size effect ($\beta_\text{body} = 0.5, \beta_\text{brain} = 2$).
+All other settings and analysis were the same as in the main text.
 
@@
   mtext('density', 4, 2.5)
-  mtext('effect brain size', 1, 2.5)
+  mtext('effect relative brain size', 1, 2.5)
   plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
@@
   mtext('density', 4, 2.5)
-  mtext('effect brain size', 1, 2.5)
+  mtext('effect relative brain size', 1, 2.5)
 
@@
 ```
+
+```{r predictor unequal effect, include=FALSE}
+# Run through magnitudes
+for(i in 1:2){
+
+  # Settings
+  set.seed(1)
+  N = 100
+  brain_effect = c(0.5, 2)[i]
+  body_effect = c(2, 0.5)[i]
+  model_predictor = cmdstan_model(path_predictor_stan)
+  model_sem = cmdstan_model(path_sem_stan)
+
+  # Simulate data
+  out = lapply(1:20, function(i){
+    body_size = rnorm(N, 0, 1)
+    rel_brain_size = rnorm(N, 0, 1)
+    brain_size = 1 * body_size + rel_brain_size
+    mu_z = body_effect * body_size + brain_effect * rel_brain_size
+    z = rnorm(N, mu_z, 1)
+
+    # Run linear model
+    linear_fit = lm(z ~ body_size + brain_size)
+
+    # Run Bayesian linear model
+    d = list(N = N, brain_size = brain_size, body_size = body_size, z = z)
+    bay_linear_fit = model_predictor$sample(data = d, 
+                                            seed = 1, 
+                                            chains = 4, 
+                                            parallel_chains = 4,
+                                            refresh = 2000)
+    post_linear_fit = bay_linear_fit$draws(format = 'draws_df')
+
+    # Run Bayesian SEM
+    bay_sem_fit = model_sem$sample(data = d, 
+                                   seed = 1, 
+                                   chains = 4, 
+                                   parallel_chains = 4,
+                                   refresh = 2000)
+    post_sem_fit = bay_sem_fit$draws(format = 'draws_df')
+
+    return(list(linear_fit = linear_fit, 
+                post_linear_fit = post_linear_fit, 
+                post_sem_fit = post_sem_fit,
+                body_size = body_size,
+                brain_size = brain_size))
+  })
+
+  # Plot the fit
+  pdf(sprintf('%s/predictor %s.pdf', path_supplemental_figures, c('strong body', 'strong brain')[i]), 9, 5)
+  par(mfrow = c(2,3), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = 'linear model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
+                                     summary(out[[i]]$linear_fit)$coefficients[3,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect brain size', 1, 2.5)
+  mtext('brain', 2, 4, font = 2, las = 2, adj = 0)
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = 'Bayesian linear model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_brain), pch = 16, cex = 3, 
+                       col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect brain size', 1, 2.5)
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = 'Bayesian SEM model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_brain_z), pch = 16, cex = 3, 
+                       col = alpha('#16A085', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect relative brain size', 1, 2.5)
+
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2],
+                                     summary(out[[i]]$linear_fit)$coefficients[2,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  mtext('body', 2, 4, font = 2, las = 2, adj = 0)
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), pch = 16, cex = 3, 
+                       col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_body_z), pch = 16, cex = 3, 
+                       col = alpha('#16A085', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  dev.off()
+
+} # end i loop
+```
+
+```{r plot predictor-strong-body, echo=FALSE, out.width = "65%", fig.cap= "Model with strong effect of body size and weak effect of brain size.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_predictor_strong_body)
+```
+
+```{r plot predictor-strong-brain, echo=FALSE, out.width = "65%", fig.cap= "Model with strong effect of brain size and weak effect of body size.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_predictor_strong_brain)
+```

### 2022-07-09 manuscript.Rmd
Adding supplemental materials as appendix instead. Also updating all references to supplemental materials.

@@
 
-To test if sample size had an effect on which model performed best, additional simulations were run and analysed for 20 and 1000 species (see supplemental results).
-To test if more or less informative priors had an effect on which model performed best, and how often the best model included the true effect in the credible interval, additional simulations were run with informative and vague priors (see supplemental materials).
+To test if sample size had an effect on which model performed best, additional simulations were run and analysed for 20 and 1000 species (see Appendix).
+To test if more or less informative priors had an effect on which model performed best, and how often the best model included the true effect in the credible interval, additional simulations were run with informative and vague priors (see Appendix).
 
@@
 
-To test how models performed in case of unqual magnitute of the effect of body size and brain size on *z* (case II), additional simulation were run (see supplemental materials) with either strong body size effect ($\beta_\text{body} = 2, \beta_\text{brain} = 0.5$) or strong brain size effect ($\beta_\text{body} = 0.5, \beta_\text{brain} = 2$).
+To test how models performed in case of unqual magnitute of the effect of body size and brain size on *z* (case II), additional simulation were run (see Appendix) with either strong body size effect ($\beta_\text{body} = 2, \beta_\text{brain} = 0.5$) or strong brain size effect ($\beta_\text{body} = 0.5, \beta_\text{brain} = 2$).
 
@@
 
-Results were similar for simulations with smaller and larger sample sizes (see supplemental materials, Figure S1-4).
-Results were also similar for models with vague and informative prior (see supplemental materials, Figure S5-8).
+Results were similar for simulations with smaller and larger sample sizes (see Appendix, Figure 4-7).
+Results were also similar for models with vague and informative prior (see Appendix, Figure 8-11).
 The informative priors constrained the Bayesian linear models estimate of the body size effect to positive values, but none of the posterior distributions included the true value of the body size effect.
 In other words, even if the model is run with the prior centred tightly around the correct value, it estimates the effect to be absent.
-Also when varying the magnitude of the effects of body size and brain size, the structural equation model was the only model to correctly estimate the effect of body size (see S9-10).
+Also when varying the magnitude of the effects of body size and brain size, the structural equation model was the only model to correctly estimate the effect of body size (see Appendix, Figure 12-13).
 
@@
 It was the body size effect that was biased in the linear models.
-Simulations with a strong body size effect still found a positive coefficient for this effect, but were biased towards lower values (see Figure S9).
-In simulations with a strong brain size effect the body size coefficients were even negative (see Figure S10).
+Simulations with a strong body size effect still found a positive coefficient for this effect, but were biased towards lower values (see Figure 12).
+In simulations with a strong brain size effect the body size coefficients were even negative (see Figure 13).
 
@@
 
-**Data accessibility.** All data is generated in the simulation. Code is publicly available at: https://github.com/simeonqs/On_the_use_of_relative_brain_size.
+**Data accessibility.** All data is generated in the simulation.
+Code is publicly available at: https://github.com/simeonqs/On_the_use_of_relative_brain_size.
+Upon acceptance all materials will also be made publicly available on Edmund.
 
@@
 # References
+
+::: {#refs}
+:::
+
+\newpage
+
+# Appendix
+
+```{r set-up appendix, include=FALSE}
+# Paths
+path_response_stan = 'models/response.stan' # models
+path_predictor_stan = 'models/predictor.stan'
+path_sem_stan = 'models/sem.stan'
+path_supplemental_models = 'supplemental models' # folder to supplemental models
+path_supplemental_figures = 'supplemental figures' # folder to plot supplemental figures
+path_pdf_response_20 = 'supplemental figures/response 20.pdf' # pdfs
+path_pdf_response_1000 = 'supplemental figures/response 1000.pdf'
+path_pdf_predictor_20 = 'supplemental figures/predictor 20.pdf'
+path_pdf_predictor_1000 = 'supplemental figures/predictor 1000.pdf'
+path_pdf_response_informative = 'supplemental figures/response informative.pdf'
+path_pdf_response_vague = 'supplemental figures/response vague.pdf'
+path_pdf_predictor_informative = 'supplemental figures/predictor informative.pdf'
+path_pdf_predictor_vague = 'supplemental figures/predictor vague.pdf'
+path_pdf_predictor_strong_body = 'supplemental figures/predictor strong body.pdf'
+path_pdf_predictor_strong_brain = 'supplemental figures/predictor strong brain.pdf'
+```
+
+# Supplemental methods
+
+## Sample size
+
+The main text contained results for data sets with 100 species.
+To see how the models perform with a smaller or larger dataset, I simulated data for 20 and 1000 species, with 20 simulations per case and sample size.
+
+## Priors
+
+The main text contained results for Bayesian models with slightly regularising priors.
+In emperical data sets, one can often choose more informative priors.
+To test the effect of this I analysed the dataset with 100 species with priors that only allow for positive slopes, with the mean set to the simulated value (1).
+I also restricted the priors for the intecept parameters to normal(0, 0.25).
+
+To further test the robustness of the models I analysed the data with vague priors, settings intercept and slope to normal(0, 10) and standard variation parameters to exponential(0.1) (note that the mean for the exponential distribution is the inverse of the rate, such that the mean standard deviation here is 10).
+
+## Effect magnitude
+
+The main text assumes equal magnitude for the body size and brain size effect.
+To test how the models performed when the body and brain effects on z are unequal in magnitude, I ran two additional simulations with either strong body size effect ($\beta_\text{body} = 2, \beta_\text{brain} = 0.5$) or strong brain size effect ($\beta_\text{body} = 0.5, \beta_\text{brain} = 2$).
+All other settings and analysis were the same as in the main text.
+
+\pagebreak
+
+# Supplemental results
+
+```{r response-sampsize, include=FALSE}
+for(N in c(20, 1000)){
+  # Settings
+  set.seed(1)
+  z_effect = 1
+  body_effect = 1
+  model_response = cmdstan_model(path_response_stan)
+  if(N == 20) ylim = c(0, 5) else ylim = c(0, 15)
+
+  # Simulate data 20 times
+  out = lapply(1:20, function(i){
+    body_size = rnorm(N, 1, 1)
+    z = rnorm(N, 0, 1)
+    mu_brain = body_effect * body_size + z_effect * z
+    brain_size = rnorm(N, mu_brain, 1)
+
+    # Run linear model
+    linear_fit = lm(brain_size ~ body_size + z)
+
+    # Run Bayesian linear model
+    d = list(N = N, brain_size = brain_size, body_size = body_size, z = z) # save data as list
+    bay_linear_fit = model_response$sample(data = d, # run model
+                                           seed = 1, 
+                                           chains = 4, 
+                                           parallel_chains = 4, # runs on four threads
+                                           refresh = 2000)
+    post_linear_fit = bay_linear_fit$draws(format = 'draws_df') # draw posterior samples
+
+    return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit))
+  })
+
+  # Plot the fit
+  pdf(sprintf('%s/response %s.pdf', path_supplemental_figures, N), 6, 5)
+  par(mfrow = c(2,2), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, xlab = '', yaxt = 'n', main = 'linear model')
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect z', 1, 2.5)
+  abline(v = z_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
+                                     summary(out[[i]]$linear_fit)$coefficients[3,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  mtext('z', 2, 3, font = 2, las = 2, adj = 0)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = 'Bayesian linear model')
+  abline(v = z_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_z), pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect z', 1, 2.5)
+
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2], 
+                                     summary(out[[i]]$linear_fit)$coefficients[2,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  mtext('body', 2, 3, font = 2, las = 2, adj = 0)
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), 
+                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  dev.off()
+}
+```
+
+```{r plot response-20, echo=FALSE, out.width = "45%", fig.cap= "Model with sample size 20.
+Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_response_20)
+```
+
+```{r plot response-1000, echo=FALSE, out.width = "45%", fig.cap= "Model with sample size 1000.
+Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple and green density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_response_1000)
+```
+
+```{r predictor-sampsize, include=FALSE}
+for(N in c(20, 1000)){
+
+  # Settings
+  set.seed(1)
+  brain_effect = 1
+  body_effect = 1
+  model_predictor = cmdstan_model(path_predictor_stan)
+  model_sem = cmdstan_model(path_sem_stan)
+  if(N == 20) ylim = c(0, 5) else ylim = c(0, 15)
+
+  # Simulate data
+  out = lapply(1:20, function(i){
+    body_size = rnorm(N, 0, 1)
+    rel_brain_size = rnorm(N, 0, 1)
+    brain_size = 1 * body_size + rel_brain_size
+    mu_z = body_effect * body_size + brain_effect * rel_brain_size
+    z = rnorm(N, mu_z, 1)
+
+    # Run linear model
+    linear_fit = lm(z ~ body_size + brain_size)
+
+    # Run Bayesian linear model
+    d = list(N = N, brain_size = brain_size, body_size = body_size, z = z)
+    bay_linear_fit = model_predictor$sample(data = d, 
+                                            seed = 1, 
+                                            chains = 4, 
+                                            parallel_chains = 4,
+                                            refresh = 2000)
+    post_linear_fit = bay_linear_fit$draws(format = 'draws_df')
+
+    # Run Bayesian SEM
+    bay_sem_fit = model_sem$sample(data = d, 
+                                   seed = 1, 
+                                   chains = 4, 
+                                   parallel_chains = 4,
+                                   refresh = 2000)
+    post_sem_fit = bay_sem_fit$draws(format = 'draws_df')
+
+    return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit, post_sem_fit = post_sem_fit))
+  })
+
+  # Plot the fit
+  pdf(sprintf('%s/predictor %s.pdf', path_supplemental_figures, N), 9, 5)
+  par(mfrow = c(2,3), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = 'linear model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
+                                     summary(out[[i]]$linear_fit)$coefficients[3,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect brain size', 1, 2.5)
+  mtext('brain', 2, 4, font = 2, las = 2, adj = 0)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = 'Bayesian linear model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_brain), 
+                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect relative brain size', 1, 2.5)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = 'Bayesian SEM model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_brain_z), 
+                       pch = 16, cex = 3, col = alpha('#16A085', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect brain size', 1, 2.5)
+
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2],
+                                     summary(out[[i]]$linear_fit)$coefficients[2,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  mtext('body', 2, 4, font = 2, las = 2, adj = 0)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), 
+                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_body_z), 
+                       pch = 16, cex = 3, col = alpha('#16A085', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  dev.off()
+
+}
+```
+
+```{r plot predictor-20, echo=FALSE, out.width = "65%", fig.cap= "Model with sample size 20.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model  with relative brain size as predictor variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_predictor_20)
+```
+
+```{r plot predictor-1000, echo=FALSE, out.width = "65%", fig.cap= "Model with sample size 1000.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model  with relative brain size as predictor variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple and green density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_predictor_1000)
+```
+
+```{r response-prior, include=FALSE}
+for(prior_type in c('informative', 'vague')){
+
+  # Settings
+  set.seed(1)
+  N = 20
+  z_effect = 1
+  body_effect = 1
+  model_response = cmdstan_model(sprintf('%s/response_%s.stan', path_supplemental_models, prior_type))
+  ylim = c(0, 5)
+
+  # Simulate data 20 times
+  out = lapply(1:20, function(i){
+    body_size = rnorm(N, 1, 1)
+    z = rnorm(N, 0, 1)
+    mu_brain = body_effect * body_size + z_effect * z
+    brain_size = rnorm(N, mu_brain, 1)
+
+    # Run linear model
+    linear_fit = lm(brain_size ~ body_size + z)
+
+    # Run Bayesian linear model
+    d = list(N = N, brain_size = brain_size, body_size = body_size, z = z) # save data as list
+    bay_linear_fit = model_response$sample(data = d, # run model
+                                           seed = 1, 
+                                           chains = 4, 
+                                           parallel_chains = 4, # runs on four threads
+                                           refresh = 2000)
+    post_linear_fit = bay_linear_fit$draws(format = 'draws_df') # draw posterior samples
+
+    return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit))
+  })
+
+  # Plot the fit
+  pdf(sprintf('%s/response %s.pdf', path_supplemental_figures, prior_type), 6, 5)
+  par(mfrow = c(2,2), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, xlab = '', yaxt = 'n', main = 'linear model')
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect z', 1, 2.5)
+  abline(v = z_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
+                                     summary(out[[i]]$linear_fit)$coefficients[3,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  mtext('z', 2, 3, font = 2, las = 2, adj = 0)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = 'Bayesian linear model')
+  abline(v = z_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_z), pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect z', 1, 2.5)
+
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2], 
+                                     summary(out[[i]]$linear_fit)$coefficients[2,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  mtext('body', 2, 3, font = 2, las = 2, adj = 0)
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), 
+                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  dev.off()
+}
+```
+
+```{r plot response-informative, echo=FALSE, out.width = "45%", fig.cap= "Model with informative priors.
+Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_response_informative)
+```
+
+```{r plot response-vague, echo=FALSE, out.width = "45%", fig.cap= "Model with vague priors.
+Parameter estimates from the linear model and Bayesian linear model with brain size as response variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple and green density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_response_vague)
+```
+
+```{r predictor-prior, include=FALSE}
+for(prior_type in c('informative', 'vague')){
+
+  # Settings
+  set.seed(1)
+  N = 100
+  brain_effect = 1
+  body_effect = 1
+  model_predictor = cmdstan_model(sprintf('%s/predictor_%s.stan', path_supplemental_models, prior_type))
+  model_sem = cmdstan_model(sprintf('%s/sem_%s.stan', path_supplemental_models, prior_type))
+  ylim = c(0, 5)
+
+  # Simulate data
+  out = lapply(1:20, function(i){
+    body_size = rnorm(N, 0, 1)
+    rel_brain_size = rnorm(N, 0, 1)
+    brain_size = 1 * body_size + rel_brain_size
+    mu_z = body_effect * body_size + brain_effect * rel_brain_size
+    z = rnorm(N, mu_z, 1)
+
+    # Run linear model
+    linear_fit = lm(z ~ body_size + brain_size)
+
+    # Run Bayesian linear model
+    d = list(N = N, brain_size = brain_size, body_size = body_size, z = z)
+    bay_linear_fit = model_predictor$sample(data = d, 
+                                            seed = 1, 
+                                            chains = 4, 
+                                            parallel_chains = 4,
+                                            refresh = 2000)
+    post_linear_fit = bay_linear_fit$draws(format = 'draws_df')
+
+    # Run Bayesian SEM
+    bay_sem_fit = model_sem$sample(data = d, 
+                                   seed = 1, 
+                                   chains = 4, 
+                                   parallel_chains = 4,
+                                   refresh = 2000)
+    post_sem_fit = bay_sem_fit$draws(format = 'draws_df')
+
+    return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit, post_sem_fit = post_sem_fit))
+  })
+
+  # Plot the fit
+  pdf(sprintf('%s/predictor %s.pdf', path_supplemental_figures, prior_type), 9, 5)
+  par(mfrow = c(2,3), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = 'linear model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
+                                     summary(out[[i]]$linear_fit)$coefficients[3,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect brain size', 1, 2.5)
+  mtext('brain', 2, 4, font = 2, las = 2, adj = 0)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = 'Bayesian linear model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_brain), 
+                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect brain size', 1, 2.5)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = 'Bayesian SEM model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_brain_z), 
+                       pch = 16, cex = 3, col = alpha('#16A085', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect relative brain size', 1, 2.5)
+
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2],
+                                     summary(out[[i]]$linear_fit)$coefficients[2,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  mtext('body', 2, 4, font = 2, las = 2, adj = 0)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), 
+                       pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  plot(NULL, xlim = c(-0.5, 1.5), ylim = ylim, 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_body_z), 
+                       pch = 16, cex = 3, col = alpha('#16A085', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  dev.off()
+
+}
+```
+
+```{r plot predictor-informative, echo=FALSE, out.width = "65%", fig.cap= "Model with informative priors.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_predictor_informative)
+```
+
+```{r plot predictor-vague, echo=FALSE, out.width = "65%", fig.cap= "Model with vague priors.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple and green density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_predictor_vague)
+```
+
+```{r predictor unequal effect, include=FALSE}
+# Run through magnitudes
+for(i in 1:2){
+
+  # Settings
+  set.seed(1)
+  N = 100
+  brain_effect = c(0.5, 2)[i]
+  body_effect = c(2, 0.5)[i]
+  model_predictor = cmdstan_model(path_predictor_stan)
+  model_sem = cmdstan_model(path_sem_stan)
+
+  # Simulate data
+  out = lapply(1:20, function(i){
+    body_size = rnorm(N, 0, 1)
+    rel_brain_size = rnorm(N, 0, 1)
+    brain_size = 1 * body_size + rel_brain_size
+    mu_z = body_effect * body_size + brain_effect * rel_brain_size
+    z = rnorm(N, mu_z, 1)
+
+    # Run linear model
+    linear_fit = lm(z ~ body_size + brain_size)
+
+    # Run Bayesian linear model
+    d = list(N = N, brain_size = brain_size, body_size = body_size, z = z)
+    bay_linear_fit = model_predictor$sample(data = d, 
+                                            seed = 1, 
+                                            chains = 4, 
+                                            parallel_chains = 4,
+                                            refresh = 2000)
+    post_linear_fit = bay_linear_fit$draws(format = 'draws_df')
+
+    # Run Bayesian SEM
+    bay_sem_fit = model_sem$sample(data = d, 
+                                   seed = 1, 
+                                   chains = 4, 
+                                   parallel_chains = 4,
+                                   refresh = 2000)
+    post_sem_fit = bay_sem_fit$draws(format = 'draws_df')
+
+    return(list(linear_fit = linear_fit, 
+                post_linear_fit = post_linear_fit, 
+                post_sem_fit = post_sem_fit,
+                body_size = body_size,
+                brain_size = brain_size))
+  })
+
+  # Plot the fit
+  pdf(sprintf('%s/predictor %s.pdf', path_supplemental_figures, c('strong body', 'strong brain')[i]), 9, 5)
+  par(mfrow = c(2,3), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = 'linear model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
+                                     summary(out[[i]]$linear_fit)$coefficients[3,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect brain size', 1, 2.5)
+  mtext('brain', 2, 4, font = 2, las = 2, adj = 0)
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = 'Bayesian linear model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_brain), pch = 16, cex = 3, 
+                       col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect brain size', 1, 2.5)
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = 'Bayesian SEM model')
+  abline(v = brain_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_brain_z), pch = 16, cex = 3, 
+                       col = alpha('#16A085', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect relative brain size', 1, 2.5)
+
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2],
+                                     summary(out[[i]]$linear_fit)$coefficients[2,2])), 
+                       pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  mtext('body', 2, 4, font = 2, las = 2, adj = 0)
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), pch = 16, cex = 3, 
+                       col = alpha('#7D3C98', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  plot(NULL, xlim = c(-4, 4), ylim = c(0, 5), 
+       xlab = '', yaxt = 'n', main = '')
+  abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+  for(i in 1:20) lines(density(out[[i]]$post_sem_fit$b_body_z), pch = 16, cex = 3, 
+                       col = alpha('#16A085', 0.5))
+  axis(4)
+  mtext('density', 4, 2.5)
+  mtext('effect body size', 1, 2.5)
+  dev.off()
+
+} # end i loop
+```
+
+```{r plot predictor-strong-body, echo=FALSE, out.width = "65%", fig.cap= "Model with strong effect of body size and weak effect of brain size.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_predictor_strong_body)
+```
+
+```{r plot predictor-strong-brain, echo=FALSE, out.width = "65%", fig.cap= "Model with strong effect of brain size and weak effect of body size.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple density plots are the posterior distributions from the Bayesian models."}
+knitr::include_graphics(path_pdf_predictor_strong_brain)
+```
+
+# References

### 2022-07-11 manuscript.Rmd
Updating main text. Only minor changes.

@@
 This becomes clear when looking at the causal structure of such a case.
-The best way to decide which variables to include in an analysis is the use of causal graphs [@glymour2016causal, @wright1934method].
+The best way to decide which variables to include in an analysis is the use of causal graphs [@glymour2016causal; @wright1934method].
 @laubach2021biologist presented a simple guide to draw Directed Acyclic Graphs (DAG), in which all variables are connected with arrows that indicate the direction of causality.
 
-To point out the difference between the case with relative brain size as response vs predictor variable, I will present the underlying assumed causal structure presented in Figure 1.
+To point out the difference between the case with relative brain size as response vs predictor variable, I will present the underlying assumed causal structure visualised in Figure 1.
 To avoid discussion about directionality I have simply named the third variable *z*.
@@
 Then I plotted the estimated coefficient of body size and *z* to show how well parameters were retrieved.
-Additionally the bias (difference between estimate and simulated value) and coverage (proportion of times the true value was within 2SE distance from the estimate) are reported.
+Additionally I reported the bias (difference between estimate and simulated value) and coverage (proportion of times the true value was within 2SE distance from the estimate).
 
@@
 The last three lines of the model are the priors for all parameters.
-Note that the priors for the slopes ($\beta$ and $\gamma$) are set to normal(0, 1), which regularises them slightly and are centred no effect of the predictors.
+Note that the priors for the slopes ($\beta$ and $\gamma$) are set to normal(0, 1), which regularises them slightly and center them around no effect of the predictors.
 For empirical studies theory might provide more informative priors, which would further increase the accuracy of the model.
@@
 correlation = sapply(out, function(x) cor(x$body_size, x$brain_size))
-mean_correlation = mean(correlation)
+mean_correlation = round(mean(correlation), 2)
 ```
@@
 knitr::kable(table_combined, format = 'markdown', caption = 'Bias and coverage of all models.
-Mean and standard deviation bias is calculated from the distance between the point estimate and simulated value.
+The mean and standard deviation of the bias is calculated from the distance between the point estimate and simulated value.
 Coverage is the proportion of times that the simulated value was withing the point estimate +- 2SE.')
@@
 
-Several recent studies have claimed to study the effect of relative brain size by including both absolute brain size and body size as predictors [@street2017coevolution; @gonzalez2010large; @isler2009why, @maklakov2011brains, @samia2015brain].
+Several recent studies have claimed to study the effect of relative brain size by including both absolute brain size and body size as predictors [@street2017coevolution; @gonzalez2010large; @isler2009why; @maklakov2011brains; @samia2015brain].
 These studies also reported the direct effect of body size and sometimes drew conclusions based on the sign of this effect.
@@
 The use of an artificial variable *z* might be confusing, so it might be informative to give examples for both cases and explain the evolutionary causal structure.
-The social brain hypothesis is possible the most hotly debated version of case I, with relative brain size as response variable.
+The social brain hypothesis is possibly the most hotly debated version of case I, with relative brain size as response variable.
 Here brain size is a function of body size and some measure of sociality (*z*).
@@
 First brought up by @jolly1966lemur, this hypotheses has since received much attention both with residual brain size and absolute brain size as response variable (for a review see @acedo2016critical).
-The second case, where relative brain size is a predictor variable, has received less attention, but several studies have attempted to show that relative brain size can cause longevity, with relatively larger brained species living longer [@street2017coevolution, @yu2018large, @decasien2018encephalization, @jimenez2020long].
+The second case, where relative brain size is a predictor variable, has received less attention, but several studies have attempted to show that relative brain size can cause longevity, with relatively larger brained species living longer [@street2017coevolution; @yu2018large; @decasien2018encephalization; @jimenez2020long].
 The understanding is that cognitive ability causes both increased brain size and the ability to survive longer.
@@
 
-For this paper I drew all variables from normal distributions with zero mean and standard deviation one.
+For this paper I drew all independent variables from normal distributions with zero mean and standard deviation one.
 I furthermore simulated *z* as a function of relative brain size, rather than simulating both *z* and brain size as a function of the unobserved cognitive ability (as depicted in Figure 1b).
@@
 
-# Supplemental methods
-
-## Sample size
+## Supplemental methods
+
+### Sample size
 
@@
 
-## Priors
+### Priors
 
@@
 
-## Effect magnitude
+### Effect magnitude
 
@@
 
-# Supplemental results
+## Supplemental results
 
@@
 ```
-
-# References

### 2022-08-09 manuscript.Rmd
Minor revisions for Ecology and Evolution.

@@
 ---
-title: "On the use of relative brain size"
+title: "Using relative brain size as predictor variable: serious pitfalls and solutions"
 author: "Simeon Q. Smeele $^{1,2,3}$"
 output:
+  pdf_document: default
   word_document: default
-  pdf_document: default
 header-includes:
@@
 
-$^2$Department of Human Behavior, Ecology and Culture, Max Planck Institute for Evolutionary
-Anthropology, Leipzig, Germany
+$^2$Department of Human Behavior, Ecology and Culture, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany
 
@@
 
-**Author for correspondence:** Simeon Q. Smeele, Max Planck Institute of Animal Behavior, Am Obstberg 1, 78315 Radolfzell am Bodensee, ssmeele@ab.mpg.de
+**Author for correspondence:** Simeon Q. Smeele, Max Planck Institute of Animal Behavior, Am Obstberg 1, 78315 Radolfzell am Bodensee, [ssmeele\@ab.mpg.de](mailto:ssmeele@ab.mpg.de){.email}
 
@@
 1.
-There is a long standing interest in the effect of relative brain size on other life history variables in a comparative context.
+There is a long-standing interest in the effect of relative brain size on other life history variables in a comparative context.
 Historically, residuals have been used to calculate these effects, but more recently it has been recognised that regression on residuals is not good practice.
@@
 2.
-I use a simple simulation to illustrate how a case with brain size as response variable differs from a case with relative brain size as predictor variable.
+I use a simple simulation to illustrate how a case in which brain size is a response variable differs from a case in which relative brain size is a predictor variable.
 I use the simulated data to test which modelling approach can estimate the underlying causal effects for each case.
@@
 More recently another statistical caveat has been highlighted (for an overview see @freckleton2002misuse).
-The use of residuals leads to biased estimates if correlation exists between the predictor variables, as is often the case in the comparative literature.
+The use of residuals leads to biased estimates if a correlation exists between the predictor variables, as is often the case in the comparative literature.
 The proposed solution is to include body size in a multiple regression, which is very similar to the use of residual brain size, if - and only if - relative brain size is the response variable.
@@
 To avoid discussion about directionality I have simply named the third variable *z*.
+This might be confusing, so I will give an example for both cases and explain the evolutionary causal structure.
+In the first case brain size is the response variable.
+The social brain hypothesis is possibly the most hotly debated version of this case.
+Here brain size is a function of body size and some measure of sociality (*z*).
+The idea is that social species experience evolutionary pressure to increase brain size to deal with complex social situations.
+First brought up by @jolly1966lemur, this hypothesis has since received much attention both with residual brain size and absolute brain size as response variable (for a review see @acedo2016critical).
+The second case, where relative brain size is a predictor variable, has received less attention, but several studies have attempted to show that relative brain size can cause longevity, with relatively larger brained species living longer [@street2017coevolution; @yu2018large; @decasien2018encephalization; @jimenez2020long].
+The understanding is that cognitive ability causes both increased brain size and the ability to survive longer.
+This might be unintuitive, since mechanistically cognition is caused by the brain, but in an evolutionary context it makes sense, since species that require more cognitive flexibility experience evolutionary pressure to increase brain size.
+Body size also causes an increase in longevity through several processes (e.g., reduced predation risk, lower metabolism).
+
 In the first case both body size and *z* cause brain size (see Figure 1a).
@@
 However, body size explains most of the variation in brain size.
-Therefore it is still a good idea to include this variable to get a more precise estimate of the effect of *z* on brain size.
+Therefore, it is still a good idea to include this variable to get a more precise estimate of the effect of *z* on brain size.
 Or in other words on relative brain size, since allometry is now accounted for.
@@
 The second case is causally more opaque (see Figure 1b).
-In general the question of interest is how cognitive ability influences the third variable *z*.
-But since this cannot be measured reliably across species (therefore denoted by a U for unobserved), relative brain size is used as a proxy.
+In general, the question of interest is how cognitive ability influences the third variable *z*.
+But since cognitive ability cannot be measured reliably across species (therefore denoted by a U for unobserved), relative brain size is used as a proxy.
 This is where it becomes tricky.
@@
 This is problematic, because the estimated effect of brain size on *z* now also contains some of the effect of body size on *z*.
-Therefore the estimated effect of body size on *z* will be biased.
+Therefore, the estimated effect of body size on *z* will be biased.
 There is no way to correct for this using the variables in the DAG in a multiple regression.
@@
 ```{r dag, echo=FALSE, out.width = "100%", fig.cap= "Directed Acyclic Graph of the two cases.
-Arrows indicatethe direction of causality.
+Arrows indicate the direction of causality.
 z stands for the third variable of interest.
@@
 b) Case II: relative brain size is a proxy predictor variable.
-U stands for the unobserved variable cognitive ability."}
+U stands for the unobserved cognitive ability."}
 knitr::include_graphics(path_pdf_DAG)
@@
 
-To show the difference between a case where relative brain size is the *response* variable and where it is a *predictor* variable I simulated two simple cases with three variables.
-To simplify the simulation I have not attempted to simulate realistic values for body size and relative brain size, but just used values with mean = 0 and standard deviation = 1.
+To show the difference between a case where relative brain size is the *response* variable and where it is a *predictor* variable I simulated two simple cases with three variables (all code is publicly available at <https://github.com/simeonqs/Using_relative_brain_size_as_predictor_variable-_serious_pitfalls_and_solutions>).
+To simplify the simulation, I have not attempted to simulate realistic values for body size and relative brain size, but just used values with mean = 0 and standard deviation = 1.
 This allows me to draw general conclusions that are not sensitive to the scale of the variables.
 It should also be noted that body and brain size are normally measured on the natural scale, but log-transformed before analysis.
-The asumption is that causal effects are linear on the logarithmic scale.
+The assumption is that causal effects are linear on the log-log scale.
 I have therefore simulated both body and brain size to be on the logarithmic scale.
@@
 Then I plotted the estimated coefficient of body size and *z* to show how well parameters were retrieved.
-Additionally I reported the bias (difference between estimate and simulated value) and coverage (proportion of times the true value was within 2SE distance from the estimate).
+Additionally, I reported the bias (difference between estimate and simulated value) and coverage (proportion of times the true value was within 2SE distance from the frequentist estimates and 2SD from the Bayesian estimates).
+
+To test how models performed with an additional path from body size to *z*, I ran an additional simulation with this causal structure (see Appendix).
 
@@
 I analysed the resulting data with both a frequentist and Bayesian linear model where brain size and body size were included as predictor variables.
-Additionally I analysed the data with a Bayesian structural equation model that included sub-models for all causal paths:
+Additionally, I analysed the data with a Bayesian structural equation model that included sub-models for all causal paths:
 
@@
 The last three lines of the model are the priors for all parameters.
-Note that the priors for the slopes ($\beta$ and $\gamma$) are set to normal(0, 1), which regularises them slightly and center them around no effect of the predictors.
+Note that the priors for the slopes ($\beta$ and $\gamma$) are set to normal(0, 1), which regularises them slightly and centre them around no effect of the predictors.
 For empirical studies theory might provide more informative priors, which would further increase the accuracy of the model.
 
-To test how models performed in case of unqual magnitute of the effect of body size and brain size on *z* (case II), additional simulation were run (see Appendix) with either strong body size effect ($\beta_\text{body} = 2, \beta_\text{brain} = 0.5$) or strong brain size effect ($\beta_\text{body} = 0.5, \beta_\text{brain} = 2$).
-
-To visualise the results I plotted all parameter estimates of the main model (for *z*) and reported the bias (difference between estimate and simulated value) and coverage (proportion of times the true value was within 2SE distance from the estimate).
+To test how models performed in case of unequal magnitude of the effect of body size and brain size on *z* (case II), additional simulations were run (see Appendix) with either strong body size effect ($\beta_\text{body} = 2, \beta_\text{brain} = 0.5$) or strong brain size effect ($\beta_\text{body} = 0.5, \beta_\text{brain} = 2$).
+
+To visualise the results, I plotted all parameter estimates of the main model (for *z*) and reported the bias (difference between estimate and simulated value) and coverage (proportion of times the true value was within 2SE distance from the frequentist estimates and 2SD from the Bayesian estimates).
 
@@
 
-For Case I, where relative brain size was the response variable, both models estimated the parameters very well (see Figure 2, Table 1).
+For Case I, where relative brain size was the response variable, both models estimated the parameters very well (see Figure 2, Table 1), also with an additional path from body size to *z* (see Appendix, Figure 14).
 For Case II, where relative brain size was a predictor variable, the effect of brain size was estimated well by all models, but the effect of body size was only estimated correctly by the structural equation model (see Figure 3, Table 1).
@@
                sum(coverage_response_z_linear)/length(coverage_response_z_linear),
-               sum(coverage_response_z_linear_bayes)/length(coverage_response_z_linear_bayes)))
+               sum(coverage_response_z_linear_bayes)/length(coverage_response_z_linear_bayes)),
+  n = rep(20, 4))
 table_response[,4:6] = round(table_response[,4:6], 2)
@@
                sum(coverage_predictor_brain_linear_bayes)/length(coverage_predictor_brain_linear_bayes),
-               sum(coverage_predictor_brain_sem)/length(coverage_predictor_brain_sem)))
+               sum(coverage_predictor_brain_sem)/length(coverage_predictor_brain_sem)),
+  n = rep(20, 6))
 table_predictor[,4:6] = round(table_predictor[,4:6], 2)
@@
 The mean and standard deviation of the bias is calculated from the distance between the point estimate and simulated value.
-Coverage is the proportion of times that the simulated value was withing the point estimate +- 2SE.')
+Coverage is the proportion of times that the simulated value was withing the point estimate +- 2SE/SD.
+N is number of replicates.')
 ```
@@
 
-Several recent studies have claimed to study the effect of relative brain size by including both absolute brain size and body size as predictors [@street2017coevolution; @gonzalez2010large; @isler2009why; @maklakov2011brains; @samia2015brain].
+Several recent studies have claimed to study the effect of relative brain size by including both absolute brain size and body size as predictors [@street2017coevolution; @gonzalez2010large; @isler2009why; @maklakov2011brains; @samia2015brain; @sol2022neuron].
 These studies also reported the direct effect of body size and sometimes drew conclusions based on the sign of this effect.
@@
 It was the body size effect that was biased in the linear models.
-Simulations with a strong body size effect still found a positive coefficient for this effect, but were biased towards lower values (see Figure 12).
-In simulations with a strong brain size effect the body size coefficients were even negative (see Figure 13).
+Simulations with a strong body size effect still found a positive coefficient for this effect, but were biased towards lower values (see Appendix, Figure 12).
+In simulations with a strong brain size effect the body size coefficients were even negative (see Appendix, Figure 13).
 
@@
 In other words, there are two paths from relative brain size (or the unobserved cognitive ability from the DAG) to *z*: one path is direct and one path is a back-door path through the allometric effect of body size.
-In another case, where body size itself does not have an effect on *z*, it does not actually need to be included at all.
+In another case, where body size itself does not have an effect on *z*, it does not actually need to be included at all [@walmsley2022causation].
 In this simple example, using absolute brain size would be fine.
@@
 A potentially even more powerful approach was put forward by @smaers2009inferring and involves reconstructing the ancestral states of both body size and brain size.
-Recently this approach was used to test brain evolution in mammals and it was shown that relatively large brains can be achieved by highly divergent paths [@smaers2021evolution].
+Recently this approach was used to model brain evolution in mammals and it was shown that relatively large brains can be achieved by divergent paths [@smaers2021evolution].
 The authors contest the notion that there is a universal scaling between body and brain size, and instead propose that relatively large brains can just as well be a result of selection towards smaller bodies.
@@
 First, the largest non-cognitive effects were due to a shift in the slope of the body to brain relation.
-Such a shift would be less likely to affect results when considering a lower order taxonomic group (e.g., only primates) and could be partially accounted for by including multiple slopes (e.g., one per family).
+Such a shift would be less likely to affect results when considering a lower-order taxonomic group (e.g., only primates) and could be partially accounted for by including multiple slopes (e.g., one per family).
 Second, running a Bayesian multi-regime OU modelling approach is not straightforward and becomes really difficult when many covariates are included, since no step-by-step guide or R package exists.
 A compromise is to first study the allometric patterns in the taxon of interest and then decide if relative brain size can be used as proxy for cognitive ability.
-
-The use of an artificial variable *z* might be confusing, so it might be informative to give examples for both cases and explain the evolutionary causal structure.
-The social brain hypothesis is possibly the most hotly debated version of case I, with relative brain size as response variable.
-Here brain size is a function of body size and some measure of sociality (*z*).
-The idea is that social species experience evolutionary pressure to increase brain size to deal with complex social situations.
-First brought up by @jolly1966lemur, this hypotheses has since received much attention both with residual brain size and absolute brain size as response variable (for a review see @acedo2016critical).
-The second case, where relative brain size is a predictor variable, has received less attention, but several studies have attempted to show that relative brain size can cause longevity, with relatively larger brained species living longer [@street2017coevolution; @yu2018large; @decasien2018encephalization; @jimenez2020long].
-The understanding is that cognitive ability causes both increased brain size and the ability to survive longer.
-This might be unintuitive, since mechanistically cognition is caused by the brain, but in an evolutionary context it makes sense, since species that require more cognitive flexibility experience evolutionary pressure to increase brain size.
 
@@
 
-Another problem with the use of residuals was brought up by @rogell2020controlling.
-They show that many studies have oposite signs between the model with absolute brain size and the model with residual brain size.
-They argue that coefficients from linear models are difficult to interpret in a biologically meaningfull way, when body size and brain size are strongly correlated.
-In the current study, the average correlation between body size and brain size was `r mean_correlation`, which is not as strong as in some emperical studies, but still cleary problematic for the linear models.
-The SEM estimates did, however, not suffer under these conditions.
-
-Structural equation models allow for the inclusion of measurement error, imputation of missing data and phylogenetic coveriance.
+Another potential problem with the use of residuals was brought up by @rogell2020controlling.
+They show that many studies have opposite signs between the model with absolute brain size and the model with residual brain size.
+They argue that coefficients from linear models are difficult to interpret in a biologically meaningful way, when body size and brain size are strongly correlated.
+In contrasts, @walmsley2022causation show that collinearity does not cause bias per se, and that sign reversal between simple and multiple regressions are simply due to the fact that these answer different biological questions.
+They do, however, point out that estimating the effect of an agent (*z*) on brain size using multiple regression can be biased if body size is to some extend caused by brain size.
+As a preliminary solution they propose a Methods of Moments estimator and show that this approach correctly estimates the coefficients in the simulated data.
+In the current study, the average correlation between body size and brain size was `r mean_correlation`, which is not as strong as in some empirical studies, but still clearly problematic for the linear models.
+The SEM estimates did not suffer under these conditions.
+Adding additional correlation in case I did not bias the parameter estimates of the multiple regressions, but it did increase uncertainty slightly (see Appendix, Figure 14).
+
+Structural equation models allow for the inclusion of measurement error, imputation of missing data and phylogenetic covariance.
 @smeele2022coevolution published their R script for such an applied structural equation model, which can be modified to fit most comparative questions.
@@
 Prior information about parameters can be used to increase accuracy.
-E.g., it is highly unlikely that an increase in body size would directly lead to a decrease in body size.
-The prior for this relationship should therefore be set to only include positive values.
+E.g., it is highly unlikely that an increase in body size would directly lead to a decrease in brain size.
+The prior for this relationship could therefore be set to only include positive values.
 This creates a clear advantage over frequentist linear models, which allow any real number and can therefore produce scientifically nonsensical results.
@@
 
-The aim of this paper was to show how a case where relative brain size is a response variable differs from a case where it is a predictor variable.
-From a simple simulation it becomes clear that one cannot include absolute brain size and body size as covariates if relative brain size is theorised to predict the third variable.
+The aim of this paper was to show that multiple regressions fail to estimate causal effects when relative brain size is a predictor variable.
 Multiple other approaches can be used instead and the design of the model should be based on the assumed causal structure of the system.
@@
 **Data accessibility.** All data is generated in the simulation.
-Code is publicly available at: https://github.com/simeonqs/On_the_use_of_relative_brain_size.
-Upon acceptance all materials will also be made publicly available on Edmund.
+Code is publicly available at: <https://github.com/simeonqs/Using_relative_brain_size_as_predictor_variable-_serious_pitfalls_and_solutions>.
+Upon acceptance all materials will also be made publicly available on Edmond.
 
@@
 path_pdf_predictor_strong_brain = 'supplemental figures/predictor strong brain.pdf'
+path_pdf_response_iii = 'supplemental figures/response case iii.pdf'
 ```
@@
 To test how the models performed when the body and brain effects on z are unequal in magnitude, I ran two additional simulations with either strong body size effect ($\beta_\text{body} = 2, \beta_\text{brain} = 0.5$) or strong brain size effect ($\beta_\text{body} = 0.5, \beta_\text{brain} = 2$).
+All other settings and analysis were the same as in the main text.
+
+### Additional causal path
+
+To test how the multiple regressions can deal with additional causal complexity I simulated a case with brain size as the response variable (similar to case I), but with an additional causal path from body size to *z*.
 All other settings and analysis were the same as in the main text.
@@
 ```
+
+```{r response-case-iii, include=FALSE}
+# Settings
+set.seed(1)
+N = 100
+z_effect_brain = 1
+z_effect_body = 1
+body_effect = 1
+model_response = cmdstan_model(path_response_stan)
+
+# Simulate data 20 times
+out = lapply(1:20, function(i){
+  z = rnorm(N, 0, 1)
+  mu_body = z_effect_body * z
+  body_size = rnorm(N, mu_body, 1)
+  mu_brain = body_effect * body_size + z_effect_brain * z
+  brain_size = rnorm(N, mu_brain, 1)
+
+  # Run linear model
+  linear_fit = lm(brain_size ~ body_size + z)
+
+  # Run Bayesian linear model
+  d = list(N = N, brain_size = brain_size, body_size = body_size, z = z) # save data as list
+  bay_linear_fit = model_response$sample(data = d, # run model
+                                         seed = 1, 
+                                         chains = 4, 
+                                         parallel_chains = 4, # runs on four threads
+                                         refresh = 2000)
+  post_linear_fit = bay_linear_fit$draws(format = 'draws_df') # draw posterior samples
+
+  return(list(linear_fit = linear_fit, post_linear_fit = post_linear_fit))
+})
+
+# Plot the fit
+pdf(path_pdf_response_iii, 7, 5.5)
+par(mfrow = c(2,2), oma = c(1, 4, 1, 1), mar = c(4, 1, 2, 3))
+plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), xlab = '', yaxt = 'n', main = 'linear model')
+axis(4)
+mtext('density', 4, 2.5)
+mtext('effect z', 1, 2.5)
+abline(v = z_effect_brain, lty = 2, lwd = 5, col = alpha('black', 0.5))
+for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[3],
+                                   summary(out[[i]]$linear_fit)$coefficients[3,2])), 
+                     pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+mtext('z', 2, 3, font = 2, las = 2, adj = 0)
+plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
+     xlab = '', yaxt = 'n', main = 'Bayesian linear model')
+abline(v = z_effect_brain, lty = 2, lwd = 5, col = alpha('black', 0.5))
+for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_z), pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
+axis(4)
+mtext('density', 4, 2.5)
+mtext('effect z', 1, 2.5)
+
+plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
+     xlab = '', yaxt = 'n', main = '')
+abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+for(i in 1:20) lines(density(rnorm(1e5, out[[i]]$linear_fit$coefficients[2], 
+                                   summary(out[[i]]$linear_fit)$coefficients[2,2])), 
+                     pch = 16, cex = 3, col = alpha('#D68910', 0.5))
+mtext('body', 2, 3, font = 2, las = 2, adj = 0)
+axis(4)
+mtext('density', 4, 2.5)
+mtext('effect body size', 1, 2.5)
+plot(NULL, xlim = c(-0.5, 1.5), ylim = c(0, 5), 
+     xlab = '', yaxt = 'n', main = '')
+abline(v = body_effect, lty = 2, lwd = 5, col = alpha('black', 0.5))
+for(i in 1:20) lines(density(out[[i]]$post_linear_fit$b_body), pch = 16, cex = 3, col = alpha('#7D3C98', 0.5))
+axis(4)
+mtext('density', 4, 2.5)
+mtext('effect body size', 1, 2.5)
+dev.off()
+```
+
+```{r plot response-iii, echo=FALSE, out.width = "65%", fig.cap= "Model with relative brain size as response variable and an additional causal path from body size to z.
+Parameter estimates from the linear model and Bayesian linear model.
+Dashed grey line is the true value.
+Orange density plots are normal distributions based on the mean and SE from the linear model.
+Purple density plots are the posterior distributions from the Bayesian model."}
+knitr::include_graphics(path_pdf_response_iii)
+```

### 2022-08-22 manuscript.Rmd
Minor updates for final version for Ecology and Evolution. Adding the DOI for Edmond repo.

@@
 output:
+  word_document: default
   pdf_document: default
-  word_document: default
 header-includes:
@@
 To control for allometric scaling (larger species have larger brains), brain size is measured relative to body size.
-There exists a large diversity of relative brain size measures [@healy2007critique], but some decades ago the use of residual brain size was proposed and quickly became the most popular [@jerison1973evolution].
+There exists a large diversity of relative brain size measures [@healy2007critique], but some decades ago the use of residual brain size was proposed and quickly became the most popular measure [@jerison1973evolution].
 This measure is the residual from a model that has log body size as predictor and log brain size as response variable.
@@
 Results were similar for simulations with smaller and larger sample sizes (see Appendix, Figure 4-7).
-Results were also similar for models with vague and informative prior (see Appendix, Figure 8-11).
+Results were also similar for models with vague and informative priors (see Appendix, Figure 8-11).
 The informative priors constrained the Bayesian linear models estimate of the body size effect to positive values, but none of the posterior distributions included the true value of the body size effect.
@@
 
-```{r plot predictor, echo=FALSE, out.width = "100%", fig.cap= "Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
+```{r plot predictor, echo=FALSE, out.width = "100%", fig.cap= "Parameter estimates from the linear model, Bayesian linear model and Bayesian structural equation model with relative brain size as predictor variable.
 Dashed grey line is the true value.
@@
 Prior information about parameters can be used to increase accuracy.
-E.g., it is highly unlikely that an increase in body size would directly lead to a decrease in brain size.
+For example, it is highly unlikely that an increase in body size would directly lead to a decrease in brain size.
 The prior for this relationship could therefore be set to only include positive values.
@@
 **Data accessibility.** All data is generated in the simulation.
-Code is publicly available at: <https://github.com/simeonqs/Using_relative_brain_size_as_predictor_variable-_serious_pitfalls_and_solutions>.
-Upon acceptance all materials will also be made publicly available on Edmond.
+Code is publicly available on Github: <https://github.com/simeonqs/Using_relative_brain_size_as_predictor_variable-_serious_pitfalls_and_solutions>.
+It is also available on Edmond: <https://doi.org/10.17617/3.PXZF2T>.
 
@@
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple and green density plots are the posterior distributions from the Bayesian models."}
+Purple density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_response_1000)
@@
 ```{r plot predictor-20, echo=FALSE, out.width = "65%", fig.cap= "Model with sample size 20.
-Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model  with relative brain size as predictor variable.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian structural equation model with relative brain size as predictor variable.
 Dashed grey line is the true value.
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple density plots are the posterior distributions from the Bayesian models."}
+Purple and green density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_predictor_20)
@@
 ```{r plot predictor-1000, echo=FALSE, out.width = "65%", fig.cap= "Model with sample size 1000.
-Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model  with relative brain size as predictor variable.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian structural equation model with relative brain size as predictor variable.
 Dashed grey line is the true value.
@@
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple and green density plots are the posterior distributions from the Bayesian models."}
+Purple density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_response_vague)
@@
 ```{r plot predictor-informative, echo=FALSE, out.width = "65%", fig.cap= "Model with informative priors.
-Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian structural equation model with relative brain size as predictor variable.
 Dashed grey line is the true value.
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple density plots are the posterior distributions from the Bayesian models."}
+Purple and green density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_predictor_informative)
@@
 ```{r plot predictor-vague, echo=FALSE, out.width = "65%", fig.cap= "Model with vague priors.
-Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian structural equation model with relative brain size as predictor variable.
 Dashed grey line is the true value.
@@
 ```{r plot predictor-strong-body, echo=FALSE, out.width = "65%", fig.cap= "Model with strong effect of body size and weak effect of brain size.
-Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian structural equation model with relative brain size as predictor variable.
 Dashed grey line is the true value.
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple density plots are the posterior distributions from the Bayesian models."}
+Purple and green density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_predictor_strong_body)
@@
 ```{r plot predictor-strong-brain, echo=FALSE, out.width = "65%", fig.cap= "Model with strong effect of brain size and weak effect of body size.
-Parameter estimates from the linear model, Bayesian linear model and Bayesian strucutral equation model with relative brain size as predictor variable.
+Parameter estimates from the linear model, Bayesian linear model and Bayesian structural equation model with relative brain size as predictor variable.
 Dashed grey line is the true value.
 Orange density plots are normal distributions based on the mean and SE from the linear model.
-Purple density plots are the posterior distributions from the Bayesian models."}
+Purple and green density plots are the posterior distributions from the Bayesian models."}
 knitr::include_graphics(path_pdf_predictor_strong_brain)
